### 2019-03-04 draft_charts.Rmd
Add data and drafts

---
output:
  html_document:
    number_sections: FALSE
    self_contained: TRUE
    code_folding: hide
    toc: TRUE
    toc_float: TRUE
    css: www/web_report.css
    editor_options:
      chunk_output_type: console
editor_options:
  chunk_output_type: console
---

<link rel="stylesheet" href="//fonts.googleapis.com/css?family=Lato" />

![](www/images/urban-institute-logo.png)

```{r rmarkdown-setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r setup}
options(scipen = 999)

library(tidyverse)
library(urbnthemes)

set_urbn_defaults(style = "print")
```

### N Charts About Future Retirement Security plots
### Aaron R. Williams
### `r format(Sys.time(), "%B %d, %Y %H:%M %p")`

## Chart 1: The share of retirees who will likely experience a decline in living standards will grow over time

* Numerator: age 70, price-adjusted, per-capita annuitized income with imputed rent and other income
* Denominator: mean price-adjusted, per-capita earnings from ages 50 to 59

```{r chart1}
cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")

read_csv("data/figure1.csv") %>%
  filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
  filter(doby %in% cohorts) %>%
  group_by(doby) %>%
  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
  ggplot(aes(doby, rep_rate_threshold)) +
  geom_col() +
  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(title = "Age 70 Percentage of People Who Fail to Replace 75% of Age 50-59",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()   
```

## Chart 2: Retirement Income at Age 70 is Rising

```{r chart2}
read_csv("data/figure2.csv") %>%
  filter(doby %in% cohorts) %>%
  ggplot(aes(doby, ncharts_aftertax_income_P50)) +
  geom_col() +
  geom_text(mapping = aes(label = scales::dollar(ncharts_aftertax_income_P50, largest_with_cents = 0)), 
            vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(title = "Age-70, per-capita, price-adjusted income",
       subtitle = "Median shared annuitized income with imputed rent, other income, and taxes",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()
```

## Chart 3 (old): The Gender Gap in Retirement Income is Narrowing

```{r chart3}
read_csv("data/figure3.csv") %>%
  filter(doby %in% cohorts) %>%
  ggplot(aes(doby, ncharts_aftertax_income_P50, fill = male)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::dollar(ncharts_aftertax_income_P50, largest_with_cents = 0)), 
            position = position_dodge(width = 0.7), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
  labs(title = "Age-70, per-capita, price-adjusted, shared annuitized income with imputed rent, other income, and taxes",
       subtitle = "By sex",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis() 
```

## Chart 3 (new): The Gender Gap in Social Secuirty Benefits is Narrowing

```{r, chart3b}
read_csv("data/figure3b.csv") %>%
  filter(doby %in% cohorts) %>%
  ggplot(aes(doby, Ownssb_P50, fill = male)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::dollar(Ownssb_P50, largest_with_cents = 0)), 
            position = position_dodge(width = 0.7), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
  labs(title = "Age-70 price-adjusted Social Security Benefits by Sex",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis() 
```

## Chart 4: Women's Lifetime Earnings are Increasing

```{r chart4}
read_csv("data/figure4.csv") %>%
  filter(doby %in% cohorts) %>%
  filter(!is.na(male)) %>%
  ggplot(aes(doby, lifetime_earn_P50, fill = male)) +
  geom_col(position = "dodge") +
  geom_text(mapping = aes(label = scales::dollar(lifetime_earn_P50, largest_with_cents = 0)), 
            position = position_dodge(width = 0.7), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
  labs(title = "Price-adjusted earnings from ages 25 to 69 at age 70",
       subtitle = "By sex",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis() 
```

## Chart 5: More Women Are Earnings Worker's Social Security Benefits

```{r chart5}
benefits <- c("Worker only", "Worker supplemented by spousal or survivor", "Spouse or survivor only")

read_csv("data/figure5.csv") %>%
  filter(doby %in% cohorts) %>%
  select(doby,
         `Worker supplemented by spousal or survivor` = dualben_Mean,
         `Worker only` = wrkonly_Mean,
         `Spouse or survivor only` = auxonly_Mean) %>%
  gather(key = "benefit_type", value = "share", -doby) %>%
  mutate(benefit_type = factor(benefit_type, levels = benefits)) %>%
  ggplot(aes(benefit_type, share, fill = doby)) +
  geom_col(position = "dodge") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), labels = scales::percent_format(accuracy = 1)) +
  labs(title = "Social Security Benefit Type at Age 70 for Female Beneficiaries (%)",
       x = NULL,
       y = NULL) +
  remove_ticks()
```

## Chart 6: Racial and ethnic disparities in retirement income will narrow modestly

* We are concerned about showing this because of how racial/ethnic differences in income are estimated and projected. We don’t have a lot of certainty about how gaps will change in the future.

```{r chart6}
read_csv("data/figure6.csv") %>%
  filter(doby %in% cohorts) %>%
  filter(!is.na(raceeth)) %>%
  ggplot(aes(doby, ncharts_aftertax_income_P50, fill = raceeth)) +
  geom_col(position = "dodge") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
                     labels = scales::dollar) +
  labs(title = "Age-70, per-capita, price-adjusted, shared annuitized income\nwith imputed rent, other income, and taxes",
       subtitle = "By race and ethnicity",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks()
```

Count

```{r chart6-count}
read_csv("data/figure6ns.csv") %>%
  filter(doby %in% cohorts) %>%
  filter(!is.na(raceeth)) %>%
  ggplot(aes(doby, count_N, fill = raceeth)) +
  geom_col(position = "dodge") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
                     labels = scales::comma) +
  labs(title = "Counts: Age-70, per-capita, price-adjusted, shared annuitized\nwith imputed rent, other income, and taxes",
       subtitle = "By race and ethnicity",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks()
```

## Chart 7: Social Security Remains a Key Pillar of Retirement Security

* Quntiles are based on per-capita, price-adjusted, shared annuitized income with imputed rent, other income, federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70. The quintiles aren't just based on people at age 70. 
* Values are the mean of the different sources of income in each quintile.

```{r chart7}
incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")

read_csv("data/figure7.csv") %>%
  filter(doby %in% cohorts) %>%
  mutate(`Financial income` = pcint_Mean + pcdiv_Mean + pcrentinc_Mean + pcdeltaIRA_Mean, 
         `Other income` = pcNotModel_Mean) %>%
  select(nchartQuin,
         doby,
         Earnings = PCearn_Mean, 
         `Social Security` = PCssb_Mean, 
         `DB pension income` = PCdbinc_Mean, 
         `Financial income`, `Other income`) %>%
  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
  mutate(income_type = factor(income_type, levels = incomes)) %>%
  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
  geom_col(position = "fill") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
                     labels = scales::percent) +
  facet_wrap(~doby) +
  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
       x = "Income quintile",
       y = NULL) +
  remove_ticks()  
```

## Chart 8: Social Security’s Financing Problems Pose Additional Risks

* Among people with Social Security benefits

```{r chart8}
read_csv("data/figure8.csv") %>%
  filter(doby %in% cohorts) %>%
  select(doby,
         `Scheduled law` = PCssb_P50,
         `Payable benefits` = PCssb_payable_P50) %>%
  gather(key = "Benefit type", value = "benefit", -doby) %>%
  mutate(`Benefit type` = factor(`Benefit type`, levels = c("Scheduled law", "Payable benefits"))) %>%
  ggplot(aes(doby, benefit, fill = `Benefit type`)) +
  geom_col(position = "dodge") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
                     labels = scales::dollar) +
  labs(title = "Median Annual Social Security Benefit at Age 70",
       subtitle = "Among People with Social Security Benefits",
       x = "Income quintile",
       y = NULL) +
  remove_ticks()
```

## Chart 9: The share of retirees who will likely experience a decline in living standards will grow slowly over time

With payable Social Security benefits

```{r, chart9}
read_csv("data/figure9.csv") %>%
filter(ncharts_reprate_payable %in% c("0-50%", "50-75%")) %>%
  filter(doby %in% cohorts) %>%
  group_by(doby) %>%
  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
  ggplot(aes(doby, rep_rate_threshold)) +
  geom_col() +
  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
  labs(title = "Age 70 Percentage of People Who fail to Replace 75% of Age 50-59 Earnings",
       subtitle = "Social Security benefits under payable law",
       x = "Birth cohort",
       y = NULL) +
  remove_ticks() +
  remove_axis()
```

### 2019-03-07 draft_charts.Rmd
Fix chart 7 data. Add more replacement rates. Add more breakdowns for chart 7. Add chart 10.

@@
 
-## Chart 1: The share of retirees who will likely experience a decline in living standards will grow over time
+## Chart 1: The share of retirees who will likely experience a decline in living standards will grow over time {.tabset}
 
@@
 * Denominator: mean price-adjusted, per-capita earnings from ages 50 to 59
+
+### 75% replacement rate
 
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who Fail to Replace 75% of Age 50-59",
+  labs(title = "Age 70 Percentage of People Who Fail to Replace 75% of Age 50-59 Earnings",
+       x = "Birth cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()   
+```
+
+### 80% replacement rate
+
+```{r chart1b}
+cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
+
+read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%")) %>%
+  filter(doby %in% cohorts) %>%
+  group_by(doby) %>%
+  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
+  ggplot(aes(doby, rep_rate_threshold)) +
+  geom_col() +
+  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
+  labs(title = "Age 70 Percentage of People Who Fail to Replace 80% of Age 50-59 Earnings",
+       x = "Birth cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()   
+```
+
+### 85% replacement rate
+
+```{r chart1c}
+cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
+
+read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%")) %>%
+  filter(doby %in% cohorts) %>%
+  group_by(doby) %>%
+  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
+  ggplot(aes(doby, rep_rate_threshold)) +
+  geom_col() +
+  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
+  labs(title = "Age 70 Percentage of People Who Fail to Replace 85% of Age 50-59 Earnings",
+       x = "Birth cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()   
+```
+
+### 90% replacement rate
+
+```{r chart1d}
+cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
+
+read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%")) %>%
+  filter(doby %in% cohorts) %>%
+  group_by(doby) %>%
+  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
+  ggplot(aes(doby, rep_rate_threshold)) +
+  geom_col() +
+  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
+  labs(title = "Age 70 Percentage of People Who Fail to Replace 90% of Age 50-59 Earnings",
        x = "Birth cohort",
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Age-70 price-adjusted Social Security Benefits by Sex",
+  labs(title = "Median Age-70 Price-Adjusted Social Security Benefits by Sex",
        x = "Birth cohort",
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Price-adjusted earnings from ages 25 to 69 at age 70",
+  labs(title = "Median Price-adjusted Earnings from Ages 25 to 69 at Age 70",
        subtitle = "By sex",
@@
   filter(!is.na(raceeth)) %>%
+  filter(raceeth != "4.Other") %>%
   ggplot(aes(doby, ncharts_aftertax_income_P50, fill = raceeth)) +
@@
 
-Count
+## Chart 7: Social Security Remains a Key Pillar of Retirement Security {.tabset}
 
-```{r chart6-count}
-read_csv("data/figure6ns.csv") %>%
+* Quntiles are based on per-capita, price-adjusted, shared annuitized income with imputed rent, other income, federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
+* Values are the mean of the different sources of income in each quintile.
+
+Financial income = per-capita interest + per-capita dividends + per-capita rental income + per-capita IRA withdrawals
+Other income = pcNotModel_Mean + per-capita imputed rent + per-capita other income
+
+### All, No Imputed Rent
+
+```{r chart7-no-imputed-rent}
+incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
+
+read_csv("data/figure7.csv") %>%
   filter(doby %in% cohorts) %>%
-  filter(!is.na(raceeth)) %>%
-  ggplot(aes(doby, count_N, fill = raceeth)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
-                     labels = scales::comma) +
-  labs(title = "Counts: Age-70, per-capita, price-adjusted, shared annuitized\nwith imputed rent, other income, and taxes",
-       subtitle = "By race and ethnicity",
-       x = "Birth cohort",
+  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
+         `Other income` = pcNotModel_Mean + pcimpr_Mean) %>%
+  select(nchartQuin,
+         doby,
+         Earnings = PCearn_Mean, 
+         `Social Security` = PCssb_Mean, 
+         `DB pension income` = PCdbinc_Mean, 
+         `Financial income`, `Other income`) %>%
+  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
+  mutate(income_type = factor(income_type, levels = incomes)) %>%
+  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
+  geom_col(position = "fill") +
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
+                     labels = scales::percent) +
+  facet_wrap(~doby) +
+  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
+       x = "Income quintile",
        y = NULL) +
@@
 
-## Chart 7: Social Security Remains a Key Pillar of Retirement Security
+### All, Imputed Rent
 
-* Quntiles are based on per-capita, price-adjusted, shared annuitized income with imputed rent, other income, federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
-The quintiles aren't just based on people at age 70.
-* Values are the mean of the different sources of income in each quintile.
-
-```{r chart7}
+```{r chart7-imputed-rent}
 incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
@@
   filter(doby %in% cohorts) %>%
-  mutate(`Financial income` = pcint_Mean + pcdiv_Mean + pcrentinc_Mean + pcdeltaIRA_Mean, 
-         `Other income` = pcNotModel_Mean) %>%
+  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
+         `Other income` = pcNotModel_Mean + pcimpr_Mean + pcimpr_Mean) %>%
+  select(nchartQuin,
+         doby,
+         Earnings = PCearn_Mean, 
+         `Social Security` = PCssb_Mean, 
+         `DB pension income` = PCdbinc_Mean, 
+         `Financial income`, `Other income`) %>%
+  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
+  mutate(income_type = factor(income_type, levels = incomes)) %>%
+  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
+  geom_col(position = "fill") +
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
+                     labels = scales::percent) +
+  facet_wrap(~doby) +
+  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
+       x = "Income quintile",
+       y = NULL) +
+  remove_ticks()  
+```
+
+### SSB > 0, No Imputed Rent
+
+**Note:** Quintiles are based on the total age 70 population.
+
+```{r chart7-ssb-no-imputed-rent}
+incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
+
+read_csv("data/figure7b.csv") %>%
+  filter(doby %in% cohorts) %>%
+  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
+         `Other income` = pcNotModel_Mean + pcimpr_Mean) %>%
+  select(nchartQuin,
+         doby,
+         Earnings = PCearn_Mean, 
+         `Social Security` = PCssb_Mean, 
+         `DB pension income` = PCdbinc_Mean, 
+         `Financial income`, `Other income`) %>%
+  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
+  mutate(income_type = factor(income_type, levels = incomes)) %>%
+  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
+  geom_col(position = "fill") +
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
+                     labels = scales::percent) +
+  facet_wrap(~doby) +
+  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
+       x = "Income quintile",
+       y = NULL) +
+  remove_ticks()  
+```
+
+### SSB > 0, Imputed Rent
+
+**Note:** Quintiles are based on the total age 70 population.
+
+```{r chart7-ssb-imputed-rent}
+incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
+
+read_csv("data/figure7b.csv") %>%
+  filter(doby %in% cohorts) %>%
+  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
+         `Other income` = pcNotModel_Mean + pcimpr_Mean + pcimpr_Mean) %>%
   select(nchartQuin,
@@
   geom_col(position = "dodge") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
+  geom_text(mapping = aes(label = scales::dollar(benefit, largest_with_cents = 0)), 
+            vjust = -1,
+             position = position_dodge(width = 0.8)) +   
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
                      labels = scales::dollar) +
@@
        y = NULL) +
-  remove_ticks()
+  remove_ticks() +
+  remove_axis()   
 ```
@@
 ```
+
+## Chart 10 (experimental): More wives are out-earning their husband's
+
+```{r}
+read_csv("data/figure10.csv") %>%
+  filter(doby <= 1985) %>%
+  mutate(doby = case_when(
+    doby <= 1945 ~ "1936-1945",
+    doby <= 1955 ~ "1946-1955",
+    doby <= 1965 ~ "1956-1965",
+    doby <= 1975 ~ "1966-1975",
+    doby <= 1985 ~ "1976-1985")
+  ) %>% 
+  ggplot(aes(x = lifetime_earn, y = slifetime_earn)) +
+  geom_point(alpha = 0.05) +
+  geom_smooth(method = "lm", formula = y ~ x - 1) +
+  scale_x_continuous(expand = c(0, 0.05), 
+                     limits = c(0, 10000000),
+                     breaks = c(0, 5000000, 10000000),
+                     labels = scales::dollar) +
+  scale_y_continuous(expand = c(0, 0.05), 
+                     limits = c(0, 10000000),
+                     breaks = c(0, 5000000, 10000000),
+                     labels = scales::dollar) +
+  facet_wrap(~doby) +
+  labs(
+    title = "Wive's lifetime earnings vs. husband's lifetime earnings",
+    subtitle = "Amongst married couples",
+    x = "Wife's lifetime earnings",
+    y = "Husband's lifetime earnings"
+  ) +
+  scatter_grid()
+
+```

### 2019-03-15 draft_charts.Rmd
Update charts outline.

@@
   labs(
-    title = "Wive's lifetime earnings vs. husband's lifetime earnings",
+    title = "Wife's lifetime earnings vs. husband's lifetime earnings",
     subtitle = "Amongst married couples",

### 2019-03-19 draft_charts.Rmd
Chnages before meeting with Alice.

@@
 knitr::opts_chunk$set(message = FALSE)
+
+knitr::opts_chunk$set(fig.width = 6.5, fig.height = 5)
 ```
@@
 library(urbnthemes)
+library(grid)
+library(gridExtra)
+library(extrafont)
+
+source("urban_style.R")
 
@@
 
-read_csv("data/figure1.csv") %>%
+chart1 <- read_csv("data/figure1.csv") %>%
   filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who Fail to Replace 75% of Age 50-59 Earnings",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
@@
   remove_axis()
+
+plotr(titler("Age 70 Percentage of People Who Fail to Replace 75% of Age 50-59 Earnings"),
+      chart1,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(2, 30, 1, 1))
 ```
@@
 ```{r chart2}
-read_csv("data/figure2.csv") %>%
+chart2 <- read_csv("data/figure2.csv") %>%
   filter(doby %in% cohorts) %>%
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age-70, per-capita, price-adjusted income",
-       subtitle = "Median shared annuitized income with imputed rent, other income, and taxes",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
@@
   remove_axis()
+
+chart2 <- plotr(titler("Age-70, per-capita, price-adjusted income"),
+      subtitler("Median shared annuitized income with imputed rent, other income, and taxes"),
+      chart2,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(1.8, 1, 30, 1, 1))
+
+ggsave("charts/chart2.pdf", 
+       chart2,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```
@@
 ```{r chart3}
-read_csv("data/figure3.csv") %>%
+chart3 <- read_csv("data/figure3.csv") %>%
   filter(doby %in% cohorts) %>%
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Age-70, per-capita, price-adjusted, shared annuitized income with imputed rent, other income, and taxes",
-       subtitle = "By sex",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
@@
   remove_axis()
+
+plotr(titler("Age-70, per-capita, price-adjusted, shared annuitized income with imputed rent, other income, and taxes"),
+      subtitler("By sex"),
+      chart3,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(2, 2, 30, 1, 1))
 ```
@@
 ```{r, chart3b}
-read_csv("data/figure3b.csv") %>%
+chart3b <- read_csv("data/figure3b.csv") %>%
   filter(doby %in% cohorts) %>%
@@
   geom_text(mapping = aes(label = scales::dollar(Ownssb_P50, largest_with_cents = 0)), 
-            position = position_dodge(width = 0.7), vjust = -1) +    
+            position = position_dodge(width = 0.75), vjust = -1) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Median Age-70 Price-Adjusted Social Security Benefits by Sex",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
@@
   remove_axis()
+
+chart3 <- plotr(titler("Median Age-70 Price-Adjusted Social Security Benefits by Sex"),
+      chart3b,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(2, 30, 1, 1))
+
+ggsave("charts/chart3.pdf", 
+       chart3,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```
@@
 ```{r chart4}
-read_csv("data/figure4.csv") %>%
+chart4 <- read_csv("data/figure4.csv") %>%
   filter(doby %in% cohorts) %>%
@@
   geom_text(mapping = aes(label = scales::dollar(lifetime_earn_P50, largest_with_cents = 0)), 
-            position = position_dodge(width = 0.7), vjust = -1) +    
+            position = position_dodge(width = 0.9), vjust = -1) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Median Price-adjusted Earnings from Ages 25 to 69 at Age 70",
-       subtitle = "By sex",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
@@
   remove_axis()
+
+chart4 <- plotr(titler("Median Price-adjusted Earnings from Ages 25 to 69 at Age 70"),
+      subtitler("By sex"),
+      chart4,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(2, 2, 30, 1, 1))
+
+ggsave("charts/chart4.pdf", 
+       chart4,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```
@@
 ```{r chart5}
-benefits <- c("Worker only", "Worker supplemented by spousal or survivor", "Spouse or survivor only")
+benefits <- c("Worker only", "Worker supplemented by\nspousal or survivor", "Spouse or survivor only")
 
-read_csv("data/figure5.csv") %>%
+chart5 <- read_csv("data/figure5.csv") %>%
   filter(doby %in% cohorts) %>%
@@
   gather(key = "benefit_type", value = "share", -doby) %>%
+  mutate(benefit_type = recode(benefit_type, 
+                               `Worker supplemented by spousal or survivor` = "Worker supplemented by\nspousal or survivor")) %>%  
   mutate(benefit_type = factor(benefit_type, levels = benefits)) %>%
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), labels = scales::percent_format(accuracy = 1)) +
-  labs(title = "Social Security Benefit Type at Age 70 for Female Beneficiaries (%)",
-       x = NULL,
+  labs(x = NULL,
        y = NULL) +
   remove_ticks()
+
+chart5 <- plotr(titler("Social Security Benefit Type at Age 70 for Female Beneficiaries (%)"),
+      chart5,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(2, 30, 1, 1))
+
+ggsave("charts/chart5.pdf", 
+       chart5,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```
@@
 ```{r chart6}
-read_csv("data/figure6.csv") %>%
+chart6 <- read_csv("data/figure6.csv") %>%
   filter(doby %in% cohorts) %>%
@@
                      labels = scales::dollar) +
-  labs(title = "Age-70, per-capita, price-adjusted, shared annuitized income\nwith imputed rent, other income, and taxes",
-       subtitle = "By race and ethnicity",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
   remove_ticks()
+
+chart6 <- plotr(titler("Age-70, per-capita, price-adjusted, shared annuitized income\nwith imputed rent, other income, and taxes"),
+      subtitler("By race and ethnicity"),
+      chart6,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(5, 1, 30, 1, 1))
+
+ggsave("charts/chart6.pdf", 
+       chart6,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```
@@
 ```{r chart8}
-read_csv("data/figure8.csv") %>%
+chart8 <- read_csv("data/figure8.csv") %>%
   filter(doby %in% cohorts) %>%
@@
                      labels = scales::dollar) +
-  labs(title = "Median Annual Social Security Benefit at Age 70",
-       subtitle = "Among People with Social Security Benefits",
-       x = "Income quintile",
+  labs(x = "Income quintile",
        y = NULL) +
@@
   remove_axis()
+
+chart8 <- plotr(titler("Median Annual Social Security Benefit at Age 70"),
+      subtitler("Among People with Social Security Benefits"),
+      chart8,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(1.8, 1, 30, 1, 1))
+
+ggsave("charts/chart8.pdf", 
+       chart8,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```
@@
 ```{r, chart9}
-read_csv("data/figure9.csv") %>%
+chart9 <- read_csv("data/figure9.csv") %>%
 filter(ncharts_reprate_payable %in% c("0-50%", "50-75%")) %>%
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who fail to Replace 75% of Age 50-59 Earnings",
-       subtitle = "Social Security benefits under payable law",
-       x = "Birth cohort",
+  labs(x = "Birth cohort",
        y = NULL) +
@@
   remove_axis()
+
+chart9 <- plotr(titler("Age 70 Percentage of People Who fail to Replace 75% of Age 50-59 Earnings"),
+      subtitler("Social Security benefits under payable law"),
+      chart9,
+      caption,
+      sourcer("DYNASIM4"),
+      heights = c(1.8, 1, 30, 1, 1))
+
+ggsave("charts/chart9.pdf", 
+       chart9,
+       device = cairo_pdf(),
+       width = 6.5,
+       height = 5,
+       units = "in")
 ```

### 2019-03-19 urban_style.R

@@ -1,0 +1,52 @@
+titler <- function(title) {
+  textGrob(title, 
+           x = unit(0, "npc"), 
+           hjust = 0, 
+           vjust = 0,
+           gp = gpar(fontsize = 12, fontfamily = "Lato"))
+}
+
+subtitler <- function(subtitle) {
+  textGrob(subtitle, 
+           x = unit(0, "npc"), 
+           hjust = 0, 
+           vjust = 0,
+           gp = gpar(fontsize = 9.5, fontfamily = "Lato"))
+}
+
+sourcer <- function(source) {
+  grobTree(
+    textGrob("Source: ", 
+             name = "source1",
+             x = unit(0, "npc"), 
+             hjust = 0, 
+             vjust = 0,
+             gp = gpar(fontsize = 8, fontfamily = "Lato", fontface = "bold")),
+    textGrob(source, 
+             x = unit(0, "npc") + grobWidth("source1"), 
+             hjust = 0, 
+             vjust = 0,
+             gp = gpar(fontsize = 8, fontfamily = "Lato"))
+  )
+}
+
+caption <- grobTree(
+  gp = gpar(fontsize = 7, hjust = 1), 
+  textGrob(label = "I N S T I T U T E", 
+           name = "caption1",
+           x = unit(1, "npc"),  
+           y = unit(0, "npc"),
+           hjust = 1, 
+           vjust = 0),
+  textGrob(label = "U R B A N  ", 
+           x = unit(1, "npc") - grobWidth("caption1") - unit(0.01, "lines"),         
+           y = unit(0, "npc"), 
+           hjust = 1, 
+           vjust = 0, 
+           gp = gpar(col = "#1696d2"))
+)
+
+plotr <- function(..., heights) {
+  grid.arrange(...,
+               heights = heights) 
+}

### 2019-04-30 draft_charts.Rmd
Update before sending for review in mid-April.

--- 2019-03-15 draft_charts.Rmd
@@
       chunk_output_type: console
+self_contained: no      
 editor_options:
@@
 knitr::opts_chunk$set(message = FALSE)
+
+knitr::opts_chunk$set(fig.width = 6.5, fig.height = 5, fig.path = "figures/", dev = c("png", "cairo_pdf"))
 ```
@@
 
-* Numerator: age 70, price-adjusted, per-capita annuitized income with imputed rent and other income
+* Numerator: age 70, price-adjusted, per-capita annuitized income
 * Denominator: mean price-adjusted, per-capita earnings from ages 50 to 59
+
+```{r}
+cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
+
+#' rename_cohort
+#'
+#' @param x a vector with cohorts
+#'
+#' @return a vector of renamed cohorts
+rename_cohort <- function(x) {
+  cohort <- list("1936-1945" = "Pre-boomers (1936–45)", 
+                 "1946-1955" = "Early boomers (1946–55)", 
+                 "1956-1965" = "Late boomers (1956–65)", 
+                 "1966-1975" = "Gen Xers (1966–75)", 
+                 "1976-1985" = "Early Millenials (1976-1985)")
+
+  if (is.null(cohort[[x]])) {
+    print("ERROR: no cohort match")
+    stop()
+  } else {
+    return(cohort[[x]])
+  }
+}
+
+rename_cohort <- function(x, two_lines = TRUE) {
+  if (two_lines == TRUE) {
+    cohort <- recode(x,
+      `1936-1945` = "Pre-boomers\n(1936–45)", 
+      `1946-1955` = "Early boomers\n(1946–55)", 
+      `1956-1965` = "Late boomers\n(1956–65)", 
+      `1966-1975` = "Gen Xers\n(1966–75)", 
+      `1976-1985` = "Early millenials\n(1976-1985)")
+
+    factor(cohort, levels = c("Pre-boomers\n(1936–45)", 
+                              "Early boomers\n(1946–55)", 
+                              "Late boomers\n(1956–65)", 
+                              "Gen Xers\n(1966–75)", 
+                              "Early millenials\n(1976-1985)"))
+  } else {
+    cohort <- recode(x,
+      `1936-1945` = "Pre-boomers (1936–45)", 
+      `1946-1955` = "Early boomers (1946–55)", 
+      `1956-1965` = "Late boomers (1956–65)", 
+      `1966-1975` = "Gen Xers (1966–75)", 
+      `1976-1985` = "Early millenials (1976-1985)")
+
+    factor(cohort, levels = c("Pre-boomers (1936–45)", 
+                              "Early boomers (1946–55)", 
+                              "Late boomers (1956–65)", 
+                              "Gen Xers (1966–75)", 
+                              "Early millenials (1976-1985)"))
+  }
+}
+
+```
 
@@
 ```{r chart1}
-cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
-
-read_csv("data/figure1.csv") %>%
+chart1 <- read_csv("data/figure1.csv") %>%
   filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
   filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%
+  group_by(doby) %>%
+  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
+  ggplot(aes(doby, rep_rate_threshold)) +
+  geom_col() +
+  geom_text(mapping = aes(label = scales::percent_format(accuracy = 0.1)(rep_rate_threshold / 100)), vjust = -1) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
+                     limits = c(0, 50)) +
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()
+
+# create the output plot
+plotr(titler("An Increasing Share of Retirees Won't Replace Pre-Retirement Incomes"),
+      subtitler("Age 70 percentage of people who fail to replace 75% of age 50-59 earnings"),
+      chart1,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(2, 2, 30, 1, 1))
+```
+
+### 80% replacement rate
+
+```{r chart1b}
+chart1 <- read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%")) %>%
+  filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%  
   group_by(doby) %>%
@@
   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who Fail to Replace 75% of Age 50-59 Earnings",
-       x = "Birth cohort",
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
+                     limits = c(0, 50)) +
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Age 70 Percentage of People Who Fail to Replace 80% of Age 50-59 Earnings"),
+      chart1,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(2, 30, 1, 1))
 ```
 
-### 80% replacement rate
+### 85% replacement rate
 
-```{r chart1b}
-cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
-
-read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%")) %>%
+```{r chart1c}
+chart1 <- read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%")) %>%
   filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%  
   group_by(doby) %>%
@@
   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who Fail to Replace 80% of Age 50-59 Earnings",
-       x = "Birth cohort",
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
+                     limits = c(0, 50)) +
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Age 70 Percentage of People Who Fail to Replace 85% of Age 50-59 Earnings"),
+      chart1,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(2, 30, 1, 1))
 ```
 
-### 85% replacement rate
+### 90% replacement rate
 
-```{r chart1c}
-cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
-
-read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%")) %>%
+```{r chart1d}
+chart1 <- read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%")) %>%
   filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%
   group_by(doby) %>%
@@
   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who Fail to Replace 85% of Age 50-59 Earnings",
-       x = "Birth cohort",
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
+                     limits = c(0, 50)) +
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Age 70 Percentage of People Who Fail to Replace 90% of Age 50-59 Earnings"),
+      chart1,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 30, 1, 1))
 ```
 
-### 90% replacement rate
+### 100% replacement rate
 
-```{r chart1d}
-cohorts <- c("1936-1945", "1946-1955", "1956-1965", "1966-1975", "1976-1985")
-
-read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%")) %>%
+```{r chart1e}
+chart1 <- read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%", "90-95%", "95-100%")) %>%
   filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%
   group_by(doby) %>%
@@
   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who Fail to Replace 90% of Age 50-59 Earnings",
-       x = "Birth cohort",
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
+                     limits = c(0, 50)) +
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Age 70 Percentage of People Who Fail to Replace 100% of Age 50-59 Earnings"),
+      chart1,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 30, 1, 1))
 ```
 
-## Chart 2: Retirement Income at Age 70 is Rising
+## Chart 2: Retirement Incomes are Increasing
 
 ```{r chart2}
-read_csv("data/figure2.csv") %>%
+chart2 <- read_csv("data/figure2.csv") %>%
   filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%  
   ggplot(aes(doby, ncharts_aftertax_income_P50)) +
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age-70, per-capita, price-adjusted income",
-       subtitle = "Median shared annuitized income with imputed rent, other income, and taxes",
-       x = "Birth cohort",
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Retirement Incomes are Increasing"),
+      subtitler("Median age-70, per-capita, price-adjusted shared annuitized income less taxes"),
+      chart2,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
 ```
 
-## Chart 3 (old): The Gender Gap in Retirement Income is Narrowing
+## Chart 3: The Gender Gap in Social Secuirty Benefits is Narrowing
 
-```{r chart3}
-read_csv("data/figure3.csv") %>%
+```{r, chart3}
+chart3b <- read_csv("data/figure3b.csv") %>%
+  mutate(male = recode(male, 
+                       `1.Male` = "Male",
+                       `2.Female` = "Female")) %>%  
   filter(doby %in% cohorts) %>%
-  ggplot(aes(doby, ncharts_aftertax_income_P50, fill = male)) +
+  mutate(doby = rename_cohort(doby)) %>%  
+  ggplot(aes(doby, Ownssb_P50, fill = male)) +
   geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::dollar(ncharts_aftertax_income_P50, largest_with_cents = 0)), 
-            position = position_dodge(width = 0.7), vjust = -1) +    
+  geom_text(mapping = aes(label = scales::dollar(Ownssb_P50, largest_with_cents = 0)), 
+            position = position_dodge(width = 0.75), vjust = -1) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Age-70, per-capita, price-adjusted, shared annuitized income with imputed rent, other income, and taxes",
-       subtitle = "By sex",
-       x = "Birth cohort",
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
-```
 
-## Chart 3 (new): The Gender Gap in Social Secuirty Benefits is Narrowing
-
-```{r, chart3b}
-read_csv("data/figure3b.csv") %>%
-  filter(doby %in% cohorts) %>%
-  ggplot(aes(doby, Ownssb_P50, fill = male)) +
-  geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::dollar(Ownssb_P50, largest_with_cents = 0)), 
-            position = position_dodge(width = 0.7), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Median Age-70 Price-Adjusted Social Security Benefits by Sex",
-       x = "Birth cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis() 
+# create the output plot
+plotr(titler("The Gender Gap in Social Security Benefits is Narrowing"),
+      subtitler("Median Age-70 Price-Adjusted Social Security Benefits by Sex"),
+      chart3b,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 2, 30, 1, 1))
 ```
@@
 ```{r chart4}
-read_csv("data/figure4.csv") %>%
+chart4 <- read_csv("data/figure4.csv") %>%
+  mutate(male = recode(male, 
+                       `1.Male` = "Male",
+                       `2.Female` = "Female")) %>%
   filter(doby %in% cohorts) %>%
   filter(!is.na(male)) %>%
+  mutate(doby = rename_cohort(doby)) %>%  
   ggplot(aes(doby, lifetime_earn_P50, fill = male)) +
@@
   geom_text(mapping = aes(label = scales::dollar(lifetime_earn_P50, largest_with_cents = 0)), 
-            position = position_dodge(width = 0.7), vjust = -1) +    
+            position = position_dodge(width = 1), vjust = -1) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(title = "Median Price-adjusted Earnings from Ages 25 to 69 at Age 70",
-       subtitle = "By sex",
-       x = "Birth cohort",
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Growth in Earnings, Especially by Women, Will Boost Future Retirement Incomes"),
+      subtitler("Median price-adjusted earnings from ages 25 to 69 at age 70 by sex"),
+      chart4,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 2, 30, 1, 1))
 ```
@@
   gather(key = "benefit_type", value = "share", -doby) %>%
+  mutate(benefit_type = recode(benefit_type, 
+                               `Worker supplemented by spousal or survivor` = "Worker supplemented by\nspousal or survivor")) %>%  
   mutate(benefit_type = factor(benefit_type, levels = benefits)) %>%
-  ggplot(aes(benefit_type, share, fill = doby)) +
+  mutate(doby = rename_cohort(doby)) %>%
+  ggplot(aes(doby, share, fill = benefit_type)) +
   geom_col(position = "dodge") +
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), labels = scales::percent_format(accuracy = 1)) +
-  labs(title = "Social Security Benefit Type at Age 70 for Female Beneficiaries (%)",
-       x = NULL,
+  labs(x = NULL,
        y = NULL) +
   remove_ticks()
+
+# create the output plot
+plotr(titler("Women’s Social Security Benefit Types are Changing"),
+      subtitler("Social Security benefit type at age 70 for female beneficiaries (%)"),
+      chart5,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 2, 30, 1, 1))
 ```
 
-## Chart 6: Racial and ethnic disparities in retirement income will narrow modestly
-
-* We are concerned about showing this because of how racial/ethnic differences in income are estimated and projected. We don’t have a lot of certainty about how gaps will change in the future.
+## Chart 6: Racial and ethnic disparities in retirement income will persist
 
 ```{r chart6}
-read_csv("data/figure6.csv") %>%
+chart6 <- read_csv("data/figure6.csv") %>%
   filter(doby %in% cohorts) %>%
@@
   filter(raceeth != "4.Other") %>%
+  mutate(raceeth = recode(raceeth, 
+                          `1.White, Non-Hisp` = "White, non-Hispanic",
+                          `2.Black, Non-Hisp` = "Black, non-Hispanic",
+                          `3.Hispanic` = "Hispanic"
+                          )) %>%  
+  mutate(doby = rename_cohort(doby)) %>%  
   ggplot(aes(doby, ncharts_aftertax_income_P50, fill = raceeth)) +
@@
                      labels = scales::dollar) +
-  labs(title = "Age-70, per-capita, price-adjusted, shared annuitized income\nwith imputed rent, other income, and taxes",
-       subtitle = "By race and ethnicity",
-       x = "Birth cohort",
+  labs(x = "Birth Cohort",
        y = NULL) +
   remove_ticks()
+
+# create the output plot
+plotr(titler("Racial and Ethnic Disparities in Retirement Income Will Persist"),
+      subtitler("Age-70, per-capita, price-adjusted, shared annuitized income less taxes by race and ethnicity"),
+      chart6,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
 ```
 
-## Chart 7: Social Security Remains a Key Pillar of Retirement Security {.tabset}
+## Chart 7: Social Security Is a Pillar of Retirement Security {.tabset}
 
-* Quntiles are based on per-capita, price-adjusted, shared annuitized income with imputed rent, other income, federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
+* Quintiles are based on per-capita, price-adjusted, shared annuitized income with federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
 * Values are the mean of the different sources of income in each quintile.
@@
 Financial income = per-capita interest + per-capita dividends + per-capita rental income + per-capita IRA withdrawals
-Other income = pcNotModel_Mean + per-capita imputed rent + per-capita other income
 
-### All, No Imputed Rent
+Other income = pcNotModel_Mean
 
-```{r chart7-no-imputed-rent}
+### Arrangement 1
+
+```{r chart7}
 incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
@@
   mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
-         `Other income` = pcNotModel_Mean + pcimpr_Mean) %>%
+         `Other income` = pcNotModel_Mean) %>%
   select(nchartQuin,
@@
   gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
+  mutate(doby = rename_cohort(doby, two_lines = FALSE)) %>% 
   mutate(income_type = factor(income_type, levels = incomes)) %>%
@@
 
-### All, Imputed Rent
+### Arrangement 2
 
-```{r chart7-imputed-rent}
+```{r chart7-alt}
 incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
 
-read_csv("data/figure7.csv") %>%
+chart7 <- read_csv("data/figure7.csv") %>%
   filter(doby %in% cohorts) %>%
   mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
-         `Other income` = pcNotModel_Mean + pcimpr_Mean + pcimpr_Mean) %>%
+         `Other income` = pcNotModel_Mean) %>%
   select(nchartQuin,
@@
   mutate(income_type = factor(income_type, levels = incomes)) %>%
-  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
+  ggplot(aes(doby, income, fill = income_type)) +
   geom_col(position = "fill") +
@@
                      labels = scales::percent) +
-  facet_wrap(~doby) +
-  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
-       x = "Income quintile",
-       y = NULL) +
-  remove_ticks()  
-```
-
-### SSB > 0, No Imputed Rent
-
-**Note:** Quintiles are based on the total age 70 population.
-
-```{r chart7-ssb-no-imputed-rent}
-incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
-
-read_csv("data/figure7b.csv") %>%
-  filter(doby %in% cohorts) %>%
-  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
-         `Other income` = pcNotModel_Mean + pcimpr_Mean) %>%
-  select(nchartQuin,
-         doby,
-         Earnings = PCearn_Mean, 
-         `Social Security` = PCssb_Mean, 
-         `DB pension income` = PCdbinc_Mean, 
-         `Financial income`, `Other income`) %>%
-  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
-  mutate(income_type = factor(income_type, levels = incomes)) %>%
-  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
-  geom_col(position = "fill") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
-                     labels = scales::percent) +
-  facet_wrap(~doby) +
-  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
-       x = "Income quintile",
-       y = NULL) +
-  remove_ticks()  
-```
-
-### SSB > 0, Imputed Rent
-
-**Note:** Quintiles are based on the total age 70 population.
-
-```{r chart7-ssb-imputed-rent}
-incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
-
-read_csv("data/figure7b.csv") %>%
-  filter(doby %in% cohorts) %>%
-  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
-         `Other income` = pcNotModel_Mean + pcimpr_Mean + pcimpr_Mean) %>%
-  select(nchartQuin,
-         doby,
-         Earnings = PCearn_Mean, 
-         `Social Security` = PCssb_Mean, 
-         `DB pension income` = PCdbinc_Mean, 
-         `Financial income`, `Other income`) %>%
-  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
-  mutate(income_type = factor(income_type, levels = incomes)) %>%
-  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
-  geom_col(position = "fill") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
-                     labels = scales::percent) +
-  facet_wrap(~doby) +
-  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
-       x = "Income quintile",
-       y = NULL) +
-  remove_ticks()  
-```
-
-## Chart 8: Social Security’s Financing Problems Pose Additional Risks
-
-* Among people with Social Security benefits
-
-```{r chart8}
-read_csv("data/figure8.csv") %>%
-  filter(doby %in% cohorts) %>%
-  select(doby,
-         `Scheduled law` = PCssb_P50,
-         `Payable benefits` = PCssb_payable_P50) %>%
-  gather(key = "Benefit type", value = "benefit", -doby) %>%
-  mutate(`Benefit type` = factor(`Benefit type`, levels = c("Scheduled law", "Payable benefits"))) %>%
-  ggplot(aes(doby, benefit, fill = `Benefit type`)) +
-  geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::dollar(benefit, largest_with_cents = 0)), 
-            vjust = -1,
-             position = position_dodge(width = 0.8)) +   
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
-                     labels = scales::dollar) +
-  labs(title = "Median Annual Social Security Benefit at Age 70",
-       subtitle = "Among People with Social Security Benefits",
-       x = "Income quintile",
+  facet_wrap(~as.character(nchartQuin)) +
+  labs(x = "Birth Cohort",
        y = NULL) +
   remove_ticks() +
-  remove_axis()   
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+# create the output plot
+plotr(titler("Social Security Is a Pillar of Retirement Security"),
+      subtitler("Major sources of income at age 70 by income quintile"),
+      chart7,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
 ```
@@
 
-With payable Social Security benefits
+```{r, chart9}
 
-```{r, chart9}
-read_csv("data/figure9.csv") %>%
-filter(ncharts_reprate_payable %in% c("0-50%", "50-75%")) %>%
+scheduled <- read_csv("data/figure1.csv") %>%
+  mutate(law = "Scheduled law")
+
+payable <- read_csv("data/figure9.csv") %>%
+  mutate(law = "Payable law") %>%
+  rename(ncharts_reprate = ncharts_reprate_payable)
+
+chart9 <- bind_rows(scheduled, payable) %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
   filter(doby %in% cohorts) %>%
-  group_by(doby) %>%
+  group_by(doby, law) %>%
   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
-  ggplot(aes(doby, rep_rate_threshold)) +
-  geom_col() +
-  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
+  ungroup() %>%
+  mutate(law = factor(law, levels = c("Scheduled law", "Payable law"))) %>%
+  mutate(doby = rename_cohort(doby)) %>%  
+  ggplot(aes(doby, rep_rate_threshold, fill = law)) +
+  geom_col(position = "dodge") +
+  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), 
+            vjust = -1,
+            position = position_dodge(width = 0.8)) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
-  labs(title = "Age 70 Percentage of People Who fail to Replace 75% of Age 50-59 Earnings",
-       subtitle = "Social Security benefits under payable law",
-       x = "Birth cohort",
+  labs(x = "Birth Cohort",
        y = NULL) +
@@
   remove_axis()
+
+# create the output plot
+plotr(titler("Social Security Insolvency Will Damage Retirement Security"),
+      subtitler("Age 70 Percentage of People Who fail to Replace 75% of Age 50-59 Earnings"),
+      chart9,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
+```
+
+## Chart 12
+
+```{r, chart12}
+chart12 <- read_csv("data/figure12.csv") %>%
+  filter(doby %in% cohorts) %>%
+  rename(`Scheduled law` = ncharts_pov_Mean, 
+         `Payable law` = ncharts_pov_payable_Mean) %>%
+  gather(key = "law", value = "poverty", -doby, -Obs) %>%
+  mutate(law = factor(law, levels = c("Scheduled law", "Payable law"))) %>%
+  ggplot(aes(doby, poverty, fill = law)) +
+  geom_col(position = "dodge") +
+  geom_text(mapping = aes(label = scales::percent_format(accuracy = 0.1)(poverty)), 
+            vjust = -1,
+            position = position_dodge(width = 0.8)) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()
+
+# create the output plot
+plotr(titler("Social Security Insolvency Will Increase Poverty"),
+      subtitler("Percentage of people in poverty at age 70"),
+      chart12,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
 ```
@@
 ```
+
+## Chart 11
+
+```{r}
+read_csv("data/figure11.csv") %>%
+  mutate(year = doby + 70,
+         insolvency_delta = (ncharts_income_payable - ncharts_income) / ncharts_income,
+         quintile = cut(lifetime_earn, 5)) %>%
+  ggplot(aes(lifetime_earn, ncharts_reprate)) +
+  geom_point(alpha = 0.01)
+
+read_csv("data/figure11.csv") %>%
+  ggplot(aes(ncharts_income, ncharts_reprate)) +
+  geom_point(alpha = 0.01) +
+  scale_x_continuous(limits = c(0, 1000000)) +
+  scale_y_continuous(limits = c(0, 10))
+
+read_csv("data/figure11.csv") %>%
+  mutate(percentile = ntile(ncharts_income, n = 100)) %>%
+  ggplot(aes(percentile, ncharts_reprate)) +
+  geom_point(alpha = 0.01) +
+  scale_x_continuous(limits = c(0, 100)) +
+  scale_y_continuous(limits = c(0, 10))
+```
+
+## Chart 13 (experimental): Failure to replace 75% of income spans the lifetime income spectrum
+
+```{r}
+figure11 <- read_csv("data/figure11.csv")
+
+figure11 %>%
+  mutate(lifetime_earn_percentile = ntile(lifetime_earn, 100)) %>%
+  group_by(lifetime_earn_percentile) %>%
+  mutate(ncharts_reprate = ncharts_reprate < 0.75) %>%
+  summarize(ncharts_reprate = mean(ncharts_reprate)) %>%
+  ggplot(aes(lifetime_earn_percentile, ncharts_reprate)) +
+  geom_point()
+```

### 2019-04-30 urban_style.R

@@ -2,7 +2,7 @@ titler <- function(title) {
   textGrob(title, 
            x = unit(0, "npc"), 
            hjust = 0, 
-           vjust = 0,
+           vjust = 0.5,
            gp = gpar(fontsize = 12, fontfamily = "Lato"))
 }
 
@@ -10,7 +10,7 @@ subtitler <- function(subtitle) {
   textGrob(subtitle, 
            x = unit(0, "npc"), 
            hjust = 0, 
-           vjust = 0,
+           vjust = 0.5,
            gp = gpar(fontsize = 9.5, fontfamily = "Lato"))
 }
 

### 2019-05-28 draft_charts.Rmd
Edits through the morning of 2019-05-28

@@
 #' @return a vector of renamed cohorts
-rename_cohort <- function(x) {
-  cohort <- list("1936-1945" = "Pre-boomers (1936–45)", 
-                 "1946-1955" = "Early boomers (1946–55)", 
-                 "1956-1965" = "Late boomers (1956–65)", 
-                 "1966-1975" = "Gen Xers (1966–75)", 
-                 "1976-1985" = "Early Millenials (1976-1985)")
+rename_cohort <- function(x, two_lines = TRUE) {
+  if (two_lines == TRUE) {
+    cohort <- recode(x,
+      `1936-1945` = "Pre-boomers\n(1936–1945)", 
+      `1946-1955` = "Early boomers\n(1946–1955)", 
+      `1956-1965` = "Late boomers\n(1956–1965)", 
+      `1966-1975` = "Gen Xers\n(1966–1975)", 
+      `1976-1985` = "Early millennials\n(1976-1985)")
 
-  if (is.null(cohort[[x]])) {
-    print("ERROR: no cohort match")
-    stop()
+    factor(cohort, levels = c("Pre-boomers\n(1936–1945)", 
+                              "Early boomers\n(1946–1955)", 
+                              "Late boomers\n(1956–1965)", 
+                              "Gen Xers\n(1966–1975)", 
+                              "Early millennials\n(1976-1985)"))
   } else {
-    return(cohort[[x]])
+    cohort <- recode(x,
+      `1936-1945` = "Pre-boomers (1936–1945)", 
+      `1946-1955` = "Early boomers (1946–1955)", 
+      `1956-1965` = "Late boomers (1956–1965)", 
+      `1966-1975` = "Gen Xers (1966–1975)", 
+      `1976-1985` = "Early millennials (1976-1985)")
+
+    factor(cohort, levels = c("Pre-boomers (1936–1945)", 
+                              "Early boomers (1946–1955)", 
+                              "Late boomers (1956–1965)", 
+                              "Gen Xers (1966–1975)", 
+                              "Early millennials (1976-1985)"))
   }
@@
 
-rename_cohort <- function(x, two_lines = TRUE) {
-  if (two_lines == TRUE) {
-    cohort <- recode(x,
-      `1936-1945` = "Pre-boomers\n(1936–45)", 
-      `1946-1955` = "Early boomers\n(1946–55)", 
-      `1956-1965` = "Late boomers\n(1956–65)", 
-      `1966-1975` = "Gen Xers\n(1966–75)", 
-      `1976-1985` = "Early millenials\n(1976-1985)")
+#' to_ordinal
+#'
+#' @param x translate numeric integers into ordinal numbers in a character vector 
+#'
+#' @return
+#' @export
+#'
+#' @examples to_ordinal(1:30)
+to_ordinal <- function(x, capitalize = FALSE, as_factor = FALSE) {
+  # case_when(x %in% 11:13 ~ paste0(x, "th"),
+  #           x %% 10 == 0 ~ paste0(x, "th"),
+  #           x %% 10 == 1 ~ paste0(x, "st"),
+  #           x %% 10 == 2 ~ paste0(x, "nd"),
+  #           x %% 10 == 3 ~ paste0(x, "rd"),
+  #           x %% 10 %in% 4:9 ~ paste0(x, "th"),
+  #           TRUE ~ "ERROR")
 
-    factor(cohort, levels = c("Pre-boomers\n(1936–45)", 
-                              "Early boomers\n(1946–55)", 
-                              "Late boomers\n(1956–65)", 
-                              "Gen Xers\n(1966–75)", 
-                              "Early millenials\n(1976-1985)"))
+  word <- case_when(x == 0 ~ "Zeroth",
+                    x == 1 ~ "Bottom",
+                    x == 2 ~ "Second",
+                    x == 3 ~ "Third",
+                    x == 4 ~ "Fourth",
+                    x == 5 ~ "Top")
+
+  if (capitalize == TRUE) {
+    tools::toTitleCase(word)
   } else {
-    cohort <- recode(x,
-      `1936-1945` = "Pre-boomers (1936–45)", 
-      `1946-1955` = "Early boomers (1946–55)", 
-      `1956-1965` = "Late boomers (1956–65)", 
-      `1966-1975` = "Gen Xers (1966–75)", 
-      `1976-1985` = "Early millenials (1976-1985)")
-
-    factor(cohort, levels = c("Pre-boomers (1936–45)", 
-                              "Early boomers (1946–55)", 
-                              "Late boomers (1956–65)", 
-                              "Gen Xers (1966–75)", 
-                              "Early millenials (1976-1985)"))
+    word
   }
 }
-
 ```
@@
   geom_col() +
-  geom_text(mapping = aes(label = scales::percent_format(accuracy = 0.1)(rep_rate_threshold / 100)), vjust = -1) +    
+  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(rep_rate_threshold / 100)), vjust = -1) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
@@
 # create the output plot
-plotr(titler("An Increasing Share of Retirees Won't Replace Pre-Retirement Incomes"),
-      subtitler("Age 70 percentage of people who fail to replace 75% of age 50-59 earnings"),
+plotr(titler("More than One-Third of Early Millennials Will Likely See Their Living Standards Fall\nwhen They Retire"),
+      subtitler("Percentage of adults unable to replace 75% of their preretirement earnings at age 70"),
       chart1,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2, 2, 30, 1, 1))
+      heights = c(2.6, 2, 30, 1, 1))
 ```
 
-### 80% replacement rate
+<!-- ### 80% replacement rate -->
 
-```{r chart1b}
-chart1 <- read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%")) %>%
-  filter(doby %in% cohorts) %>%
-  mutate(doby = rename_cohort(doby)) %>%  
-  group_by(doby) %>%
-  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
-  ggplot(aes(doby, rep_rate_threshold)) +
-  geom_col() +
-  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     limits = c(0, 50)) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis()
+<!-- ```{r chart1b} -->
+<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
+<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%")) %>% -->
+<!--   filter(doby %in% cohorts) %>% -->
+<!--   mutate(doby = rename_cohort(doby)) %>%   -->
+<!--   group_by(doby) %>% -->
+<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
+<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
+<!--   geom_col() + -->
+<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
+<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
+<!--                      limits = c(0, 50)) + -->
+<!--   labs(x = "Birth Cohort", -->
+<!--        y = NULL) + -->
+<!--   remove_ticks() + -->
+<!--   remove_axis()    -->
 
-# create the output plot
-plotr(titler("Age 70 Percentage of People Who Fail to Replace 80% of Age 50-59 Earnings"),
-      chart1,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(2, 30, 1, 1))
-```
+<!-- # create the output plot -->
+<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 80% of Age 50-59 Earnings"), -->
+<!--       chart1, -->
+<!--       caption, -->
+<!--       sourcer("DYNASIM4, ID963"), -->
+<!--       heights = c(2, 30, 1, 1)) -->
+<!-- ``` -->
 
-### 85% replacement rate
+<!-- ### 85% replacement rate -->
 
-```{r chart1c}
-chart1 <- read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%")) %>%
-  filter(doby %in% cohorts) %>%
-  mutate(doby = rename_cohort(doby)) %>%  
-  group_by(doby) %>%
-  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
-  ggplot(aes(doby, rep_rate_threshold)) +
-  geom_col() +
-  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     limits = c(0, 50)) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis()
+<!-- ```{r chart1c} -->
+<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
+<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%")) %>% -->
+<!--   filter(doby %in% cohorts) %>% -->
+<!--   mutate(doby = rename_cohort(doby)) %>%   -->
+<!--   group_by(doby) %>% -->
+<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
+<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
+<!--   geom_col() + -->
+<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
+<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
+<!--                      limits = c(0, 50)) + -->
+<!--   labs(x = "Birth Cohort", -->
+<!--        y = NULL) + -->
+<!--   remove_ticks() + -->
+<!--   remove_axis()    -->
 
-# create the output plot
-plotr(titler("Age 70 Percentage of People Who Fail to Replace 85% of Age 50-59 Earnings"),
-      chart1,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(2, 30, 1, 1))
-```
+<!-- # create the output plot -->
+<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 85% of Age 50-59 Earnings"), -->
+<!--       chart1, -->
+<!--       caption, -->
+<!--       sourcer("DYNASIM4, ID963"), -->
+<!--       heights = c(2, 30, 1, 1)) -->
+<!-- ``` -->
 
-### 90% replacement rate
+<!-- ### 90% replacement rate -->
 
-```{r chart1d}
-chart1 <- read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%")) %>%
-  filter(doby %in% cohorts) %>%
-  mutate(doby = rename_cohort(doby)) %>%
-  group_by(doby) %>%
-  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
-  ggplot(aes(doby, rep_rate_threshold)) +
-  geom_col() +
-  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     limits = c(0, 50)) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis()
+<!-- ```{r chart1d} -->
+<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
+<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%")) %>% -->
+<!--   filter(doby %in% cohorts) %>% -->
+<!--   mutate(doby = rename_cohort(doby)) %>% -->
+<!--   group_by(doby) %>% -->
+<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
+<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
+<!--   geom_col() + -->
+<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
+<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
+<!--                      limits = c(0, 50)) + -->
+<!--   labs(x = "Birth Cohort", -->
+<!--        y = NULL) + -->
+<!--   remove_ticks() + -->
+<!--   remove_axis()    -->
 
-# create the output plot
-plotr(titler("Age 70 Percentage of People Who Fail to Replace 90% of Age 50-59 Earnings"),
-      chart1,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 30, 1, 1))
-```
+<!-- # create the output plot -->
+<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 90% of Age 50-59 Earnings"), -->
+<!--       chart1, -->
+<!--       caption, -->
+<!--       sourcer("DYNASIM4, ID963"), -->
+<!--       heights = c(1.8, 30, 1, 1)) -->
+<!-- ``` -->
 
-### 100% replacement rate
+<!-- ### 100% replacement rate -->
 
-```{r chart1e}
-chart1 <- read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%", "90-95%", "95-100%")) %>%
-  filter(doby %in% cohorts) %>%
-  mutate(doby = rename_cohort(doby)) %>%
-  group_by(doby) %>%
-  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
-  ggplot(aes(doby, rep_rate_threshold)) +
-  geom_col() +
-  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     limits = c(0, 50)) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis()
+<!-- ```{r chart1e} -->
+<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
+<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%", "90-95%", "95-100%")) %>% -->
+<!--   filter(doby %in% cohorts) %>% -->
+<!--   mutate(doby = rename_cohort(doby)) %>% -->
+<!--   group_by(doby) %>% -->
+<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
+<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
+<!--   geom_col() + -->
+<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
+<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
+<!--                      limits = c(0, 50)) + -->
+<!--   labs(x = "Birth Cohort", -->
+<!--        y = NULL) + -->
+<!--   remove_ticks() + -->
+<!--   remove_axis()    -->
 
-# create the output plot
-plotr(titler("Age 70 Percentage of People Who Fail to Replace 100% of Age 50-59 Earnings"),
-      chart1,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 30, 1, 1))
-```
+<!-- # create the output plot -->
+<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 100% of Age 50-59 Earnings"), -->
+<!--       chart1, -->
+<!--       caption, -->
+<!--       sourcer("DYNASIM4, ID963"), -->
+<!--       heights = c(1.8, 30, 1, 1)) -->
+<!-- ``` -->
 
@@
   geom_col() +
-  geom_text(mapping = aes(label = scales::dollar(ncharts_aftertax_income_P50, largest_with_cents = 0)), 
+  geom_text(mapping = aes(label = scales::dollar(signif(ncharts_aftertax_income_P50, 3), largest_with_cents = 0)), 
             vjust = -1) +    
@@
 # create the output plot
-plotr(titler("Retirement Incomes are Increasing"),
-      subtitler("Median age-70, per-capita, price-adjusted shared annuitized income less taxes"),
+plotr(titler("A Typical Early Millennial Retiree Will Receive about 30 Percent More Income Than a\nTypical Retiree Born 40 Years Earlier"),
+      subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
       chart2,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 1, 30, 1, 1))
+      heights = c(2.8, 1, 30, 1, 1))
 ```
@@
   geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::dollar(Ownssb_P50, largest_with_cents = 0)), 
+  geom_text(mapping = aes(label = scales::dollar(round(Ownssb_P50, -2), largest_with_cents = 0)), 
             position = position_dodge(width = 0.75), vjust = -1) +    
@@
 # create the output plot
-plotr(titler("The Gender Gap in Social Security Benefits is Narrowing"),
-      subtitler("Median Age-70 Price-Adjusted Social Security Benefits by Sex"),
+plotr(titler("For Early Millennials, Women’s Median Age-70 Social Security Benefits Will Rise to\n85 Percent of the Median for Men"),
+      subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
       chart3b,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 2, 30, 1, 1))
+      heights = c(2.6, 2, 30, 1, 1))
 ```
@@
   geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::dollar(lifetime_earn_P50, largest_with_cents = 0)), 
+  geom_text(mapping = aes(label = scales::dollar(round(lifetime_earn_P50, -2), largest_with_cents = 0)), 
             position = position_dodge(width = 1), vjust = -1) +    
@@
 # create the output plot
-plotr(titler("Growth in Earnings, Especially by Women, Will Boost Future Retirement Incomes"),
-      subtitler("Median price-adjusted earnings from ages 25 to 69 at age 70 by sex"),
+plotr(titler("Over Her Lifetime, a Typical Early Millennial Woman Will Earn More than Twice as Much\nas Her Pre-Boomer Counterpart"),
+      subtitler("Median earnings from ages 25 to 69 (constant 2018 dollars)"),
       chart4,
@@
 # create the output plot
-plotr(titler("Racial and Ethnic Disparities in Retirement Income Will Persist"),
-      subtitler("Age-70, per-capita, price-adjusted, shared annuitized income less taxes by race and ethnicity"),
+plotr(titler("Although Retirement Incomes Are Projected to Rise for People of Color, They\nWill Continue to Lag Far Behind Incomes for Non-Hispanic Whites"),
+      subtitler("Median annual per capita after-tax family income at age 70 (thousands of constant 2018 dollars)"),
       chart6,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 1, 30, 1, 1))
+      heights = c(2.8, 1, 30, 1, 1))
 ```
 
-## Chart 7: Social Security Is a Pillar of Retirement Security {.tabset}
+## Chart 7: Social Security Is a Pillar of Retirement Security
 
@@
 
-### Arrangement 1
-
 ```{r chart7}
@@
 
-read_csv("data/figure7.csv") %>%
-  filter(doby %in% cohorts) %>%
-  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
-         `Other income` = pcNotModel_Mean) %>%
-  select(nchartQuin,
-         doby,
-         Earnings = PCearn_Mean, 
-         `Social Security` = PCssb_Mean, 
-         `DB pension income` = PCdbinc_Mean, 
-         `Financial income`, `Other income`) %>%
-  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
-  mutate(doby = rename_cohort(doby, two_lines = FALSE)) %>% 
-  mutate(income_type = factor(income_type, levels = incomes)) %>%
-  ggplot(aes(as.character(nchartQuin), income, fill = income_type)) +
-  geom_col(position = "fill") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
-                     labels = scales::percent) +
-  facet_wrap(~doby) +
-  labs(title = "Major Sources of Income at Age 70 by Income Quintile",
-       x = "Income quintile",
-       y = NULL) +
-  remove_ticks()  
-```
-
-### Arrangement 2
-
-```{r chart7-alt}
-incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
+quintile <- c("Bottom quintile", "Second quintile", "Third quintile", "Fourth quintile", "Top quintile")
 
 chart7 <- read_csv("data/figure7.csv") %>%
+  mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
+  mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
   filter(doby %in% cohorts) %>%
@@
                      labels = scales::percent) +
-  facet_wrap(~as.character(nchartQuin)) +
+  facet_wrap(~nchartQuin) +
   labs(x = "Birth Cohort",
@@
 # create the output plot
-plotr(titler("Social Security Is a Pillar of Retirement Security"),
-      subtitler("Major sources of income at age 70 by income quintile"),
+plotr(titler("Social Security Accounts for About 80 Percent of Income for Low-Income Retirees"),
+      subtitler("Percent of age-70 per capita family income from Social Security by family income quintile"),
       chart7,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
+```
+
+```{r chart7alt}
+chart7alt <- read_csv("data/figure7.csv") %>%
+  mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
+  mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
+  mutate(ss_share = PCssb_Mean / ncharts_aftertax_income_Mean) %>%
+  filter(doby %in% cohorts) %>%
+  select(nchartQuin, doby, ss_share) %>%
+  ggplot(aes(doby, ss_share, fill = nchartQuin)) +
+  geom_col(position = "dodge") +
+  geom_text(aes(label = scales::percent_format(accuracy = 1)(ss_share)),
+            position = position_dodge(width = 0.7),
+            vjust = -0.8,
+            hjust = 0.35) +
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
+                     labels = scales::percent) +
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()
+
+# create the output plot
+plotr(titler("Social Security Accounts for About 80 Percent of Income for Low-Income Retirees"),
+      subtitler("Percent of age-70 per capita family income from Social Security by family income quintile"),
+      chart7alt,
       caption,
@@
 scheduled <- read_csv("data/figure1.csv") %>%
-  mutate(law = "Scheduled law")
+  mutate(law = "Current Social Security benefits continue")
 
 payable <- read_csv("data/figure9.csv") %>%
-  mutate(law = "Payable law") %>%
+  mutate(law = "Benefits cut 25%") %>%
   rename(ncharts_reprate = ncharts_reprate_payable)
@@
   ungroup() %>%
-  mutate(law = factor(law, levels = c("Scheduled law", "Payable law"))) %>%
+  mutate(law = factor(law, levels = c("Current Social Security benefits continue", "Benefits cut 25%"))) %>%
   mutate(doby = rename_cohort(doby)) %>%  
@@
   geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), 
+  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(rep_rate_threshold / 100)), 
             vjust = -1,
@@
 # create the output plot
-plotr(titler("Social Security Insolvency Will Damage Retirement Security"),
-      subtitler("Age 70 Percentage of People Who fail to Replace 75% of Age 50-59 Earnings"),
+plotr(titler("Retirement Incomes May Fall Short for about 4 in 10 Early Millennials if Policymakers\nIgnore Social Security’s Financing Gap"),
+      subtitler("Percentage of adults unable to replace 75% of their preretirement earnings at age 70 "),
       chart9,
@@
   filter(doby %in% cohorts) %>%
-  rename(`Scheduled law` = ncharts_pov_Mean, 
-         `Payable law` = ncharts_pov_payable_Mean) %>%
+  rename(`Current Social Security benefits continue` = ncharts_pov_Mean, 
+         `Benefits cut 25%` = ncharts_pov_payable_Mean) %>%
+  mutate(doby = rename_cohort(doby)) %>% 
   gather(key = "law", value = "poverty", -doby, -Obs) %>%
-  mutate(law = factor(law, levels = c("Scheduled law", "Payable law"))) %>%
+  mutate(law = factor(law, levels = c("Current Social Security benefits continue", "Benefits cut 25%"))) %>%
   ggplot(aes(doby, poverty, fill = law)) +
@@
 # create the output plot
-plotr(titler("Social Security Insolvency Will Increase Poverty"),
-      subtitler("Percentage of people in poverty at age 70"),
+plotr(titler("A 25 Percent Social Security Benefit Cut Could Leave About One in Eight Gen X and\nEarly Millennial Retirees in Poverty"),
+      subtitler("Percentage of adults with incomes below the federal poverty level at age 70 "),
       chart12,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 1, 30, 1, 1))
+      heights = c(2.6, 1, 30, 1, 1))
 ```
@@
 
-```{r}
+```{r, chart10}
 read_csv("data/figure10.csv") %>%
@@
   ) %>% 
+  mutate(doby = rename_cohort(doby)) %>%   
   ggplot(aes(x = lifetime_earn, y = slifetime_earn)) +
@@
 
-## Chart 11
+## Chart 13 taxes
 
-```{r}
-read_csv("data/figure11.csv") %>%
-  mutate(year = doby + 70,
-         insolvency_delta = (ncharts_income_payable - ncharts_income) / ncharts_income,
-         quintile = cut(lifetime_earn, 5)) %>%
-  ggplot(aes(lifetime_earn, ncharts_reprate)) +
-  geom_point(alpha = 0.01)
+```{r, chart13}
 
-read_csv("data/figure11.csv") %>%
-  ggplot(aes(ncharts_income, ncharts_reprate)) +
-  geom_point(alpha = 0.01) +
-  scale_x_continuous(limits = c(0, 1000000)) +
-  scale_y_continuous(limits = c(0, 10))
+chart13 <- read_csv("data/figure13.csv") %>%
+  filter(doby %in% cohorts) %>%
+  mutate(avg_annual_tax_rate = (-1 * ncharts_tax_Mean) / ncharts_income_Mean) %>%
+  mutate(doby = rename_cohort(doby)) %>%    
+  filter(is.na(nchartQuin)) %>%
+  ggplot(aes(doby, avg_annual_tax_rate)) +
+  geom_col() +
+  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(avg_annual_tax_rate)), 
+            vjust = -1,
+            position = position_dodge(width = 0.8)) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1))) +
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()
 
-read_csv("data/figure11.csv") %>%
-  mutate(percentile = ntile(ncharts_income, n = 100)) %>%
-  ggplot(aes(percentile, ncharts_reprate)) +
-  geom_point(alpha = 0.01) +
-  scale_x_continuous(limits = c(0, 100)) +
-  scale_y_continuous(limits = c(0, 10))
+plotr(titler("Millennials Will Have Higher Tax Rates in Retirement than Boomers"),
+      subtitler("Average annual tax rate at age 70 (percent)"),
+      chart13,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(2.6, 1, 30, 1, 1))
 ```
 
-## Chart 13 (experimental): Failure to replace 75% of income spans the lifetime income spectrum
+<!-- ## Chart 11 -->
 
-```{r}
-figure11 <- read_csv("data/figure11.csv")
+<!-- ```{r} -->
+<!-- read_csv("data/figure11.csv") %>% -->
+<!--   mutate(year = doby + 70, -->
+<!--          insolvency_delta = (ncharts_income_payable - ncharts_income) / ncharts_income, -->
+<!--          quintile = cut(lifetime_earn, 5)) %>% -->
+<!--   ggplot(aes(lifetime_earn, ncharts_reprate)) + -->
+<!--   geom_point(alpha = 0.01) -->
 
-figure11 %>%
-  mutate(lifetime_earn_percentile = ntile(lifetime_earn, 100)) %>%
-  group_by(lifetime_earn_percentile) %>%
-  mutate(ncharts_reprate = ncharts_reprate < 0.75) %>%
-  summarize(ncharts_reprate = mean(ncharts_reprate)) %>%
-  ggplot(aes(lifetime_earn_percentile, ncharts_reprate)) +
-  geom_point()
-```
+<!-- read_csv("data/figure11.csv") %>% -->
+<!--   ggplot(aes(ncharts_income, ncharts_reprate)) + -->
+<!--   geom_point(alpha = 0.01) + -->
+<!--   scale_x_continuous(limits = c(0, 1000000)) + -->
+<!--   scale_y_continuous(limits = c(0, 10)) -->
+
+<!-- read_csv("data/figure11.csv") %>% -->
+<!--   mutate(percentile = ntile(ncharts_income, n = 100)) %>% -->
+<!--   ggplot(aes(percentile, ncharts_reprate)) + -->
+<!--   geom_point(alpha = 0.01) + -->
+<!--   scale_x_continuous(limits = c(0, 100)) + -->
+<!--   scale_y_continuous(limits = c(0, 10)) -->
+<!-- ``` -->
+
+<!-- ## Chart 13 (experimental): Failure to replace 75% of income spans the lifetime income spectrum -->
+
+<!-- ```{r} -->
+<!-- figure11 <- read_csv("data/figure11.csv") -->
+
+<!-- figure11 %>% -->
+<!--   mutate(lifetime_earn_percentile = ntile(lifetime_earn, 100)) %>% -->
+<!--   group_by(lifetime_earn_percentile) %>% -->
+<!--   mutate(ncharts_reprate = ncharts_reprate < 0.75) %>% -->
+<!--   summarize(ncharts_reprate = mean(ncharts_reprate)) %>% -->
+<!--   ggplot(aes(lifetime_earn_percentile, ncharts_reprate)) + -->
+<!--   geom_point() -->
+<!-- ``` -->

### 2019-05-28 urban_style.R

@@ -3,7 +3,10 @@ titler <- function(title) {
            x = unit(0, "npc"), 
            hjust = 0, 
            vjust = 0.5,
-           gp = gpar(fontsize = 12, fontfamily = "Lato"))
+           gp = gpar(fontsize = 12, 
+                     fontfamily = "Lato", 
+                     lineheight = 1,
+                     linewidth = 0.5))
 }
 
 subtitler <- function(subtitle) {

### 2019-06-30 draft_charts.Rmd
Update titles and change file names.

@@
 
-## Chart 1: The share of retirees who will likely experience a decline in living standards will grow over time {.tabset}
-
-* Numerator: age 70, price-adjusted, per-capita annuitized income
-* Denominator: mean price-adjusted, per-capita earnings from ages 50 to 59
-
 ```{r}
@@
       `1966-1975` = "Gen Xers\n(1966–1975)", 
-      `1976-1985` = "Early millennials\n(1976-1985)")
+      `1976-1985` = "Xennials\n(1976-1985)")
 
@@
                               "Gen Xers\n(1966–1975)", 
-                              "Early millennials\n(1976-1985)"))
+                              "Xennials\n(1976-1985)"))
   } else {
@@
       `1966-1975` = "Gen Xers (1966–1975)", 
-      `1976-1985` = "Early millennials (1976-1985)")
+      `1976-1985` = "Xennials (1976-1985)")
 
@@
                               "Gen Xers (1966–1975)", 
-                              "Early millennials (1976-1985)"))
+                              "Xennials (1976-1985)"))
   }
@@
 
-### 75% replacement rate
+## Theme 1: Retirement Incomes are Increasing
 
-```{r chart1}
-chart1 <- read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
-  filter(doby %in% cohorts) %>%
-  mutate(doby = rename_cohort(doby)) %>%
-  group_by(doby) %>%
-  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
-  ggplot(aes(doby, rep_rate_threshold)) +
-  geom_col() +
-  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(rep_rate_threshold / 100)), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     limits = c(0, 50)) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis()
-
-# create the output plot
-plotr(titler("More than One-Third of Early Millennials Will Likely See Their Living Standards Fall\nwhen They Retire"),
-      subtitler("Percentage of adults unable to replace 75% of their preretirement earnings at age 70"),
-      chart1,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(2.6, 2, 30, 1, 1))
-```
-
-<!-- ### 80% replacement rate -->
-
-<!-- ```{r chart1b} -->
-<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
-<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%")) %>% -->
-<!--   filter(doby %in% cohorts) %>% -->
-<!--   mutate(doby = rename_cohort(doby)) %>%   -->
-<!--   group_by(doby) %>% -->
-<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
-<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
-<!--   geom_col() + -->
-<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
-<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
-<!--                      limits = c(0, 50)) + -->
-<!--   labs(x = "Birth Cohort", -->
-<!--        y = NULL) + -->
-<!--   remove_ticks() + -->
-<!--   remove_axis()    -->
-
-<!-- # create the output plot -->
-<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 80% of Age 50-59 Earnings"), -->
-<!--       chart1, -->
-<!--       caption, -->
-<!--       sourcer("DYNASIM4, ID963"), -->
-<!--       heights = c(2, 30, 1, 1)) -->
-<!-- ``` -->
-
-<!-- ### 85% replacement rate -->
-
-<!-- ```{r chart1c} -->
-<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
-<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%")) %>% -->
-<!--   filter(doby %in% cohorts) %>% -->
-<!--   mutate(doby = rename_cohort(doby)) %>%   -->
-<!--   group_by(doby) %>% -->
-<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
-<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
-<!--   geom_col() + -->
-<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
-<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
-<!--                      limits = c(0, 50)) + -->
-<!--   labs(x = "Birth Cohort", -->
-<!--        y = NULL) + -->
-<!--   remove_ticks() + -->
-<!--   remove_axis()    -->
-
-<!-- # create the output plot -->
-<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 85% of Age 50-59 Earnings"), -->
-<!--       chart1, -->
-<!--       caption, -->
-<!--       sourcer("DYNASIM4, ID963"), -->
-<!--       heights = c(2, 30, 1, 1)) -->
-<!-- ``` -->
-
-<!-- ### 90% replacement rate -->
-
-<!-- ```{r chart1d} -->
-<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
-<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%")) %>% -->
-<!--   filter(doby %in% cohorts) %>% -->
-<!--   mutate(doby = rename_cohort(doby)) %>% -->
-<!--   group_by(doby) %>% -->
-<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
-<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
-<!--   geom_col() + -->
-<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
-<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
-<!--                      limits = c(0, 50)) + -->
-<!--   labs(x = "Birth Cohort", -->
-<!--        y = NULL) + -->
-<!--   remove_ticks() + -->
-<!--   remove_axis()    -->
-
-<!-- # create the output plot -->
-<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 90% of Age 50-59 Earnings"), -->
-<!--       chart1, -->
-<!--       caption, -->
-<!--       sourcer("DYNASIM4, ID963"), -->
-<!--       heights = c(1.8, 30, 1, 1)) -->
-<!-- ``` -->
-
-<!-- ### 100% replacement rate -->
-
-<!-- ```{r chart1e} -->
-<!-- chart1 <- read_csv("data/figure1.csv") %>% -->
-<!--   filter(ncharts_reprate %in% c("0-50%", "50-75%", "75-80%", "80-85%", "85-90%", "90-95%", "95-100%")) %>% -->
-<!--   filter(doby %in% cohorts) %>% -->
-<!--   mutate(doby = rename_cohort(doby)) %>% -->
-<!--   group_by(doby) %>% -->
-<!--   summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>% -->
-<!--   ggplot(aes(doby, rep_rate_threshold)) + -->
-<!--   geom_col() + -->
-<!--   geom_text(mapping = aes(label = scales::percent(rep_rate_threshold / 100)), vjust = -1) +     -->
-<!--   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), -->
-<!--                      limits = c(0, 50)) + -->
-<!--   labs(x = "Birth Cohort", -->
-<!--        y = NULL) + -->
-<!--   remove_ticks() + -->
-<!--   remove_axis()    -->
-
-<!-- # create the output plot -->
-<!-- plotr(titler("Age 70 Percentage of People Who Fail to Replace 100% of Age 50-59 Earnings"), -->
-<!--       chart1, -->
-<!--       caption, -->
-<!--       sourcer("DYNASIM4, ID963"), -->
-<!--       heights = c(1.8, 30, 1, 1)) -->
-<!-- ``` -->
-
-## Chart 2: Retirement Incomes are Increasing
-
-```{r chart2}
+```{r theme1}
 chart2 <- read_csv("data/figure2.csv") %>%
@@
 # create the output plot
-plotr(titler("A Typical Early Millennial Retiree Will Receive about 30 Percent More Income Than a\nTypical Retiree Born 40 Years Earlier"),
-      subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
+plotr(titler("A typical Xennial retiree will receive about 30 percent more net income than a typical\nretiree born 40 years earlier"),
+      subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
       chart2,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.8, 1, 30, 1, 1))
+      heights = c(2.8, 1.5, 30, 1, 1))
 ```
 
-## Chart 3: The Gender Gap in Social Secuirty Benefits is Narrowing
+## Theme 2: Women's Lifetime Earnings are Increasing
 
-```{r, chart3}
-chart3b <- read_csv("data/figure3b.csv") %>%
-  mutate(male = recode(male, 
-                       `1.Male` = "Male",
-                       `2.Female` = "Female")) %>%  
-  filter(doby %in% cohorts) %>%
-  mutate(doby = rename_cohort(doby)) %>%  
-  ggplot(aes(doby, Ownssb_P50, fill = male)) +
-  geom_col(position = "dodge") +
-  geom_text(mapping = aes(label = scales::dollar(round(Ownssb_P50, -2), largest_with_cents = 0)), 
-            position = position_dodge(width = 0.75), vjust = -1) +    
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  remove_axis()
-
-# create the output plot
-plotr(titler("For Early Millennials, Women’s Median Age-70 Social Security Benefits Will Rise to\n85 Percent of the Median for Men"),
-      subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
-      chart3b,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(2.6, 2, 30, 1, 1))
-```
-
-## Chart 4: Women's Lifetime Earnings are Increasing
-
-```{r chart4}
+```{r theme2}
 chart4 <- read_csv("data/figure4.csv") %>%
@@
 # create the output plot
-plotr(titler("Over Her Lifetime, a Typical Early Millennial Woman Will Earn More than Twice as Much\nas Her Pre-Boomer Counterpart"),
+plotr(titler("Over her lifetime, a typical Xennial woman will earn more than twice as much as her\npre-boomer counterpart"),
       subtitler("Median earnings from ages 25 to 69 (constant 2018 dollars)"),
@@
 
-## Chart 5: More Women Are Earnings Worker's Social Security Benefits
+##  Theme 3: The share of retirees who will likely experience a decline in living standards will grow over time
 
-```{r chart5}
-benefits <- c("Worker only", "Worker supplemented by\nspousal or survivor", "Spouse or survivor only")
+* Numerator: age 70, price-adjusted, per-capita annuitized income
+* Denominator: mean price-adjusted, per-capita earnings from ages 50 to 59
 
-chart5 <- read_csv("data/figure5.csv") %>%
+```{r theme3}
+chart1 <- read_csv("data/figure1.csv") %>%
+  filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
   filter(doby %in% cohorts) %>%
-  select(doby,
-         `Worker supplemented by spousal or survivor` = dualben_Mean,
-         `Worker only` = wrkonly_Mean,
-         `Spouse or survivor only` = auxonly_Mean) %>%
-  gather(key = "benefit_type", value = "share", -doby) %>%
-  mutate(benefit_type = recode(benefit_type, 
-                               `Worker supplemented by spousal or survivor` = "Worker supplemented by\nspousal or survivor")) %>%  
-  mutate(benefit_type = factor(benefit_type, levels = benefits)) %>%
   mutate(doby = rename_cohort(doby)) %>%
-  ggplot(aes(doby, share, fill = benefit_type)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), labels = scales::percent_format(accuracy = 1)) +
-  labs(x = NULL,
-       y = NULL) +
-  remove_ticks()
-
-# create the output plot
-plotr(titler("Women’s Social Security Benefit Types are Changing"),
-      subtitler("Social Security benefit type at age 70 for female beneficiaries (%)"),
-      chart5,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 2, 30, 1, 1))
-```
-
-## Chart 6: Racial and ethnic disparities in retirement income will persist
-
-```{r chart6}
-chart6 <- read_csv("data/figure6.csv") %>%
-  filter(doby %in% cohorts) %>%
-  filter(!is.na(raceeth)) %>%
-  filter(raceeth != "4.Other") %>%
-  mutate(raceeth = recode(raceeth, 
-                          `1.White, Non-Hisp` = "White, non-Hispanic",
-                          `2.Black, Non-Hisp` = "Black, non-Hispanic",
-                          `3.Hispanic` = "Hispanic"
-                          )) %>%  
-  mutate(doby = rename_cohort(doby)) %>%  
-  ggplot(aes(doby, ncharts_aftertax_income_P50, fill = raceeth)) +
-  geom_col(position = "dodge") +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
-                     labels = scales::dollar) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks()
-
-# create the output plot
-plotr(titler("Although Retirement Incomes Are Projected to Rise for People of Color, They\nWill Continue to Lag Far Behind Incomes for Non-Hispanic Whites"),
-      subtitler("Median annual per capita after-tax family income at age 70 (thousands of constant 2018 dollars)"),
-      chart6,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(2.8, 1, 30, 1, 1))
-```
-
-## Chart 7: Social Security Is a Pillar of Retirement Security
-
-* Quintiles are based on per-capita, price-adjusted, shared annuitized income with federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
-* Values are the mean of the different sources of income in each quintile.
-
-Financial income = per-capita interest + per-capita dividends + per-capita rental income + per-capita IRA withdrawals
-
-Other income = pcNotModel_Mean
-
-```{r chart7}
-incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
-
-quintile <- c("Bottom quintile", "Second quintile", "Third quintile", "Fourth quintile", "Top quintile")
-
-chart7 <- read_csv("data/figure7.csv") %>%
-  mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
-  mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
-  filter(doby %in% cohorts) %>%
-  mutate(`Financial income` = PCwlthinc_Mean + PCRetinc_Mean, 
-         `Other income` = pcNotModel_Mean) %>%
-  select(nchartQuin,
-         doby,
-         Earnings = PCearn_Mean, 
-         `Social Security` = PCssb_Mean, 
-         `DB pension income` = PCdbinc_Mean, 
-         `Financial income`, `Other income`) %>%
-  gather(key = "income_type", value = "income", Earnings, `Social Security`, `DB pension income`, `Financial income`, `Other income`) %>%
-  mutate(income_type = factor(income_type, levels = incomes)) %>%
-  ggplot(aes(doby, income, fill = income_type)) +
-  geom_col(position = "fill") +
+  group_by(doby) %>%
+  summarize(rep_rate_threshold = sum(count_PctSum_10_count)) %>%
+  ggplot(aes(doby, rep_rate_threshold)) +
+  geom_col() +
+  geom_text(mapping = aes(label = scales::percent_format(accuracy = 1)(rep_rate_threshold / 100)), vjust = -1) +    
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     labels = scales::percent) +
-  facet_wrap(~nchartQuin) +
-  labs(x = "Birth Cohort",
-       y = NULL) +
-  remove_ticks() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-# create the output plot
-plotr(titler("Social Security Accounts for About 80 Percent of Income for Low-Income Retirees"),
-      subtitler("Percent of age-70 per capita family income from Social Security by family income quintile"),
-      chart7,
-      caption,
-      sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 1, 30, 1, 1))
-```
-
-```{r chart7alt}
-chart7alt <- read_csv("data/figure7.csv") %>%
-  mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
-  mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
-  mutate(ss_share = PCssb_Mean / ncharts_aftertax_income_Mean) %>%
-  filter(doby %in% cohorts) %>%
-  select(nchartQuin, doby, ss_share) %>%
-  ggplot(aes(doby, ss_share, fill = nchartQuin)) +
-  geom_col(position = "dodge") +
-  geom_text(aes(label = scales::percent_format(accuracy = 1)(ss_share)),
-            position = position_dodge(width = 0.7),
-            vjust = -0.8,
-            hjust = 0.35) +
-  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
-                     labels = scales::percent) +
+                     limits = c(0, 50)) +
   labs(x = "Birth Cohort",
@@
 # create the output plot
-plotr(titler("Social Security Accounts for About 80 Percent of Income for Low-Income Retirees"),
-      subtitler("Percent of age-70 per capita family income from Social Security by family income quintile"),
-      chart7alt,
+plotr(titler("Nearly one-third of Xennials will see their living standards fall when they retire"),
+      subtitler("Share of adults unable to replace 75 percent of preretirement earnings at age 70"),
+      chart1,
       caption,
       sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 1, 30, 1, 1))
+      heights = c(2, 2, 30, 1, 1))
 ```
 
-## Chart 9: The share of retirees who will likely experience a decline in living standards will grow slowly over time
+## Theme 4: The share of retirees who will likely experience a decline in living standards will grow slowly over time
 
-```{r, chart9}
+```{r, theme4}
 
@@
 # create the output plot
-plotr(titler("Retirement Incomes May Fall Short for about 4 in 10 Early Millennials if Policymakers\nIgnore Social Security’s Financing Gap"),
-      subtitler("Percentage of adults unable to replace 75% of their preretirement earnings at age 70 "),
+plotr(titler("Retirement incomes may fall short for about 4 in 10 Xennials if policymakers\nignore Social Security’s financing gap"),
+      subtitler("Share of adults unable to replace 75 percent of their preretirement earnings at age 70 "),
       chart9,
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.8, 1, 30, 1, 1))
+      heights = c(3, 1.5, 30, 1, 1))
 ```
 
-## Chart 12
+## Theme 5: Social Security Is a Pillar of Retirement Security
 
-```{r, chart12}
+* Quintiles are based on per-capita, price-adjusted, shared annuitized income with federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
+* Values are the mean of the different sources of income in each quintile.
+
+Financial income = per-capita interest + per-capita dividends + per-capita rental income + per-capita IRA withdrawals
+
+Other income = pcNotModel_Mean
+
+```{r theme5}
+incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
+quintile <- c("Bottom quintile", "Second quintile", "Third quintile", "Fourth quintile", "Top quintile")
+
+chart7alt <- read_csv("data/figure7.csv") %>%
+  mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
+  mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
+  mutate(ss_share = PCssb_Mean / ncharts_aftertax_income_Mean) %>%
+  filter(doby %in% cohorts) %>%
+  select(nchartQuin, doby, ss_share) %>%
+  ggplot(aes(doby, ss_share, fill = nchartQuin)) +
+  geom_col(position = "dodge") +
+  geom_text(aes(label = scales::percent_format(accuracy = 1)(ss_share)),
+            position = position_dodge(width = 0.7),
+            vjust = -0.8,
+            hjust = 0.35) +
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)), 
+                     labels = scales::percent) +
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()
+
+# create the output plot
+plotr(titler("Social Security accounts for about 90 percent of income for low-income retirees"),
+      subtitler("Share of per capita family income at age 70 from Social Security by income quintile"),
+      chart7alt,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(1.8, 1, 30, 1, 1))
+```
+
+## Theme 6:
+
+```{r, theme6}
 chart12 <- read_csv("data/figure12.csv") %>%
@@
 # create the output plot
-plotr(titler("A 25 Percent Social Security Benefit Cut Could Leave About One in Eight Gen X and\nEarly Millennial Retirees in Poverty"),
+plotr(titler("A 25 percent Social Security benefit cut could leave about 1 in 12 Gen X and Xennial\nretirees in poverty"),
       subtitler("Percentage of adults with incomes below the federal poverty level at age 70 "),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.6, 1, 30, 1, 1))
+      heights = c(2.8, 1.5, 30, 1, 1))
 ```
 
-## Chart 10 (experimental): More wives are out-earning their husband's
+## Theme 7: Racial and ethnic disparities in retirement income will persist
 
-```{r, chart10}
-read_csv("data/figure10.csv") %>%
-  filter(doby <= 1985) %>%
-  mutate(doby = case_when(
-    doby <= 1945 ~ "1936-1945",
-    doby <= 1955 ~ "1946-1955",
-    doby <= 1965 ~ "1956-1965",
-    doby <= 1975 ~ "1966-1975",
-    doby <= 1985 ~ "1976-1985")
-  ) %>% 
+```{r theme7}
+chart6 <- read_csv("data/figure6.csv") %>%
+  filter(doby %in% cohorts) %>%
+  filter(!is.na(raceeth)) %>%
+  filter(raceeth != "4.Other") %>%
+  mutate(raceeth = recode(raceeth, 
+                          `1.White, Non-Hisp` = "White, non-Hispanic",
+                          `2.Black, Non-Hisp` = "Black, non-Hispanic",
+                          `3.Hispanic` = "Hispanic"
+                          )) %>%  
   mutate(doby = rename_cohort(doby)) %>%  
-  ggplot(aes(x = lifetime_earn, y = slifetime_earn)) +
-  geom_point(alpha = 0.05) +
-  geom_smooth(method = "lm", formula = y ~ x - 1) +
-  scale_x_continuous(expand = c(0, 0.05), 
-                     limits = c(0, 10000000),
-                     breaks = c(0, 5000000, 10000000),
+  ggplot(aes(doby, ncharts_aftertax_income_P50, fill = raceeth)) +
+  geom_col(position = "dodge") +
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15)), 
                      labels = scales::dollar) +
-  scale_y_continuous(expand = c(0, 0.05), 
-                     limits = c(0, 10000000),
-                     breaks = c(0, 5000000, 10000000),
-                     labels = scales::dollar) +
-  facet_wrap(~doby) +
-  labs(
-    title = "Wife's lifetime earnings vs. husband's lifetime earnings",
-    subtitle = "Amongst married couples",
-    x = "Wife's lifetime earnings",
-    y = "Husband's lifetime earnings"
-  ) +
-  scatter_grid()
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks()
 
+# create the output plot
+plotr(titler("Although retirement incomes are projected to rise for people of color, they\nwill continue to lag far behind incomes for non-Hispanic white people"),
+      subtitler("Median annual per capita after-tax family income at age 70 (thousands of constant 2018 dollars)"),
+      chart6,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(3, 1.5, 30, 1, 1))
 ```
 
-## Chart 13 taxes
+## Theme 8: The Gender Gap in Social Secuirty Benefits is Narrowing
 
-```{r, chart13}
+```{r, theme8}
+chart3b <- read_csv("data/figure3b.csv") %>%
+  mutate(male = recode(male, 
+                       `1.Male` = "Male",
+                       `2.Female` = "Female")) %>%  
+  filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%  
+  ggplot(aes(doby, Ownssb_P50, fill = male)) +
+  geom_col(position = "dodge") +
+  geom_text(mapping = aes(label = scales::dollar(round(Ownssb_P50, -2), largest_with_cents = 0)), 
+            position = position_dodge(width = 0.75), vjust = -1) +    
+  scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
+  labs(x = "Birth Cohort",
+       y = NULL) +
+  remove_ticks() +
+  remove_axis()
+
+# create the output plot
+plotr(titler("For Xennials, women’s median Social Security benefits at age 70 will rise to 85 percent of\nthe median for men"),
+      subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
+      chart3b,
+      caption,
+      sourcer("DYNASIM4, ID963"),
+      heights = c(2.6, 2, 30, 1, 1))
+```
+
+## Theme 9
+
+```{r, theme9}
 
@@
 
-plotr(titler("Millennials Will Have Higher Tax Rates in Retirement than Boomers"),
-      subtitler("Average annual tax rate at age 70 (percent)"),
+plotr(titler("Average tax rates for retirees will grow over time"),
+      subtitler("Average annual tax rate at age 70"),
       chart13,
@@
 ```
-
-<!-- ## Chart 11 -->
-
-<!-- ```{r} -->
-<!-- read_csv("data/figure11.csv") %>% -->
-<!--   mutate(year = doby + 70, -->
-<!--          insolvency_delta = (ncharts_income_payable - ncharts_income) / ncharts_income, -->
-<!--          quintile = cut(lifetime_earn, 5)) %>% -->
-<!--   ggplot(aes(lifetime_earn, ncharts_reprate)) + -->
-<!--   geom_point(alpha = 0.01) -->
-
-<!-- read_csv("data/figure11.csv") %>% -->
-<!--   ggplot(aes(ncharts_income, ncharts_reprate)) + -->
-<!--   geom_point(alpha = 0.01) + -->
-<!--   scale_x_continuous(limits = c(0, 1000000)) + -->
-<!--   scale_y_continuous(limits = c(0, 10)) -->
-
-<!-- read_csv("data/figure11.csv") %>% -->
-<!--   mutate(percentile = ntile(ncharts_income, n = 100)) %>% -->
-<!--   ggplot(aes(percentile, ncharts_reprate)) + -->
-<!--   geom_point(alpha = 0.01) + -->
-<!--   scale_x_continuous(limits = c(0, 100)) + -->
-<!--   scale_y_continuous(limits = c(0, 10)) -->
-<!-- ``` -->
-
-<!-- ## Chart 13 (experimental): Failure to replace 75% of income spans the lifetime income spectrum -->
-
-<!-- ```{r} -->
-<!-- figure11 <- read_csv("data/figure11.csv") -->
-
-<!-- figure11 %>% -->
-<!--   mutate(lifetime_earn_percentile = ntile(lifetime_earn, 100)) %>% -->
-<!--   group_by(lifetime_earn_percentile) %>% -->
-<!--   mutate(ncharts_reprate = ncharts_reprate < 0.75) %>% -->
-<!--   summarize(ncharts_reprate = mean(ncharts_reprate)) %>% -->
-<!--   ggplot(aes(lifetime_earn_percentile, ncharts_reprate)) + -->
-<!--   geom_point() -->
-<!-- ``` -->

### 2019-06-30 urban_style.R

@@ -14,7 +14,9 @@ subtitler <- function(subtitle) {
            x = unit(0, "npc"), 
            hjust = 0, 
            vjust = 0.5,
-           gp = gpar(fontsize = 9.5, fontfamily = "Lato"))
+           gp = gpar(fontsize = 9.5, 
+                     fontfamily = "Lato",
+                     fontface = "italic"))
 }
 
 sourcer <- function(source) {
@@ -40,13 +42,14 @@ caption <- grobTree(
            x = unit(1, "npc"),  
            y = unit(0, "npc"),
            hjust = 1, 
-           vjust = 0),
+           vjust = 0,
+           gp = gpar(fontface = "bold")),
   textGrob(label = "U R B A N  ", 
            x = unit(1, "npc") - grobWidth("caption1") - unit(0.01, "lines"),         
            y = unit(0, "npc"), 
            hjust = 1, 
            vjust = 0, 
-           gp = gpar(col = "#1696d2"))
+           gp = gpar(col = "#1696d2", fontface = "bold"))
 )
 
 plotr <- function(..., heights) {

### 2019-07-02 draft_charts.Rmd
Commit before sending to Alice.

@@
     cohort <- recode(x,
-      `1936-1945` = "Pre-boomers\n(1936–1945)", 
-      `1946-1955` = "Early boomers\n(1946–1955)", 
-      `1956-1965` = "Late boomers\n(1956–1965)", 
-      `1966-1975` = "Gen Xers\n(1966–1975)", 
-      `1976-1985` = "Xennials\n(1976-1985)")
+      `1936-1945` = "Pre-boomers\n(1936—1945)", 
+      `1946-1955` = "Early boomers\n(1946—1955)", 
+      `1956-1965` = "Late boomers\n(1956—1965)", 
+      `1966-1975` = "Gen Xers\n(1966—1975)", 
+      `1976-1985` = "Xennials\n(1976—1985)")
 
-    factor(cohort, levels = c("Pre-boomers\n(1936–1945)", 
-                              "Early boomers\n(1946–1955)", 
-                              "Late boomers\n(1956–1965)", 
-                              "Gen Xers\n(1966–1975)", 
-                              "Xennials\n(1976-1985)"))
+    factor(cohort, levels = c("Pre-boomers\n(1936—1945)", 
+                              "Early boomers\n(1946—1955)", 
+                              "Late boomers\n(1956—1965)", 
+                              "Gen Xers\n(1966—1975)", 
+                              "Xennials\n(1976—1985)"))
   } else {
     cohort <- recode(x,
-      `1936-1945` = "Pre-boomers (1936–1945)", 
-      `1946-1955` = "Early boomers (1946–1955)", 
-      `1956-1965` = "Late boomers (1956–1965)", 
-      `1966-1975` = "Gen Xers (1966–1975)", 
-      `1976-1985` = "Xennials (1976-1985)")
+      `1936-1945` = "Pre-boomers (1936—1945)", 
+      `1946-1955` = "Early boomers (1946—1955)", 
+      `1956-1965` = "Late boomers (1956—1965)", 
+      `1966-1975` = "Gen Xers (1966—1975)", 
+      `1976-1985` = "Xennials (1976—1985)")
 
-    factor(cohort, levels = c("Pre-boomers (1936–1945)", 
-                              "Early boomers (1946–1955)", 
-                              "Late boomers (1956–1965)", 
-                              "Gen Xers (1966–1975)", 
-                              "Xennials (1976-1985)"))
+    factor(cohort, levels = c("Pre-boomers (1936—1945)", 
+                              "Early boomers (1946—1955)", 
+                              "Late boomers (1956—1965)", 
+                              "Gen Xers (1966—1975)", 
+                              "Xennials (1976—1985)"))
   }
@@
 
-## Theme 1: Retirement Incomes are Increasing
+## Theme 1: Retirement incomes will continue to rise
 
@@
   mutate(doby = rename_cohort(doby)) %>%  
-  ggplot(aes(doby, ncharts_aftertax_income_P50)) +
+  ggplot(aes(doby, pc_cen_net_income_P50)) +
   geom_col() +
-  geom_text(mapping = aes(label = scales::dollar(signif(ncharts_aftertax_income_P50, 3), largest_with_cents = 0)), 
+  geom_text(mapping = aes(label = scales::dollar(signif(pc_cen_net_income_P50, 3), largest_with_cents = 0)), 
             vjust = -1) +    
@@
 
-## Theme 2: Women's Lifetime Earnings are Increasing
+```{r}
+data2 <- read_csv("data/figure2.csv") %>%
+  filter(doby %in% cohorts)
+
+data2[data2$doby == "1966-1975", "pc_cen_net_income_P50"] / data2[data2$doby == "1936-1945", "pc_cen_net_income_P50"] - 1
+
+data2[data2$doby == "1976-1985", "pc_cen_net_income_P50"] / data2[data2$doby == "1936-1945", "pc_cen_net_income_P50"] - 1
+
+```
+
+## Theme 2: Women’s increased earnings will power retirement income growth
 
@@
 
-##  Theme 3: The share of retirees who will likely experience a decline in living standards will grow over time
+```{r}
+data4 <- read_csv("data/figure4.csv") %>%
+  filter(doby %in% cohorts)
+
+data4[data4$doby == "1966-1975" & data4$male == "2.Female", "lifetime_earn_P50"] / data4[data4$doby == "1936-1945" & data4$male == "2.Female", "lifetime_earn_P50"] - 1
+
+data4[data4$doby == "1976-1985" & data4$male == "2.Female", "lifetime_earn_P50"] / data4[data4$doby == "1936-1945" & data4$male == "2.Female", "lifetime_earn_P50"] - 1
+
+data4[data4$doby == "1976-1985" & data4$male == "1.Male", "lifetime_earn_P50"] / data4[data4$doby == "1936-1945" & data4$male == "1.Male", "lifetime_earn_P50"] - 1
+
+data4[data4$doby == "1976-1985" & data4$male == "1.Male", "lifetime_earn_P50"] / data4[data4$doby == "1976-1985" & data4$male == "2.Female", "lifetime_earn_P50"] - 1
+```
+
+## Theme 3: Despite growth in retirement incomes, retirement resources will fall short for more Gen Xers and Xennials
 
@@
 chart1 <- read_csv("data/figure1.csv") %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
+  filter(ann_reprate %in% c("0-50%", "50-75%")) %>%
   filter(doby %in% cohorts) %>%
@@
   scale_y_continuous(expand = expand_scale(mult = c(0, 0.1)),
-                     limits = c(0, 50)) +
+                     limits = c(0, 40)) +
   labs(x = "Birth Cohort",
@@
 
-## Theme 4: The share of retirees who will likely experience a decline in living standards will grow slowly over time
+## Theme 4: Social Security’s financing problems threaten retirement security
 
@@
   mutate(law = "Benefits cut 25%") %>%
-  rename(ncharts_reprate = ncharts_reprate_payable)
+  rename(ann_reprate = ann_payable_reprate)
 
 chart9 <- bind_rows(scheduled, payable) %>%
-  filter(ncharts_reprate %in% c("0-50%", "50-75%")) %>%
+  filter(ann_reprate %in% c("0-50%", "50-75%")) %>%
   filter(doby %in% cohorts) %>%
@@
 
-## Theme 5: Social Security Is a Pillar of Retirement Security
+## Theme 5: Social Security cuts would disproportionately harm low-income retirees
 
@@
 
-chart7alt <- read_csv("data/figure7.csv") %>%
+chart7alt <- read_csv("data/figure7b.csv") %>%
   mutate(nchartQuin = paste(to_ordinal(nchartQuin), "quintile")) %>%
   mutate(nchartQuin = factor(nchartQuin, levels = quintile)) %>%
-  mutate(ss_share = PCssb_Mean / ncharts_aftertax_income_Mean) %>%
+  mutate(ss_share = PCssb_Mean / pc_cen_income_Mean) %>%
   filter(doby %in% cohorts) %>%
+  mutate(doby = rename_cohort(doby)) %>%    
   select(nchartQuin, doby, ss_share) %>%
@@
 # create the output plot
-plotr(titler("Social Security accounts for about 90 percent of income for low-income retirees"),
+plotr(titler("Social Security accounts for about 85 percent of income for low-income beneficiaries"),
       subtitler("Share of per capita family income at age 70 from Social Security by income quintile"),
@@
 
-## Theme 6:
+## Theme 6: Old-age poverty rates could rise if policymakers ignore Social Security’s financial problems
 
@@
   filter(doby %in% cohorts) %>%
-  rename(`Current Social Security benefits continue` = ncharts_pov_Mean, 
-         `Benefits cut 25%` = ncharts_pov_payable_Mean) %>%
+  rename(`Current Social Security benefits continue` = poverty_status_Mean, 
+         `Benefits cut 25%` = poverty_status_payable_Mean) %>%
   mutate(doby = rename_cohort(doby)) %>% 
@@
   mutate(doby = rename_cohort(doby)) %>%  
-  ggplot(aes(doby, ncharts_aftertax_income_P50, fill = raceeth)) +
+  ggplot(aes(doby, pc_cen_net_income_P50, fill = raceeth)) +
   geom_col(position = "dodge") +
@@
 plotr(titler("Although retirement incomes are projected to rise for people of color, they\nwill continue to lag far behind incomes for non-Hispanic white people"),
-      subtitler("Median annual per capita after-tax family income at age 70 (thousands of constant 2018 dollars)"),
+      subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
       chart6,
@@
 
-## Theme 8: The Gender Gap in Social Secuirty Benefits is Narrowing
+```{r}
+data6 <- read_csv("data/figure6.csv") %>%
+  filter(doby %in% cohorts)
+
+1 - data6[data6$doby == "1936-1945" & data6$raceeth == "2.Black, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"] / data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]
+
+(data6[data6$doby == "1936-1945" & data6$raceeth == "3.Hispanic" & !is.na(data6$raceeth), "pc_cen_net_income_P50"] - data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]) / data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]
+
+```
+
+## Theme 8: The gender gap in Social Security benefits will narrow
 
@@
 
-## Theme 9
+```{r}
+data3b <- read_csv("data/figure3b.csv") %>%
+  filter(doby %in% cohorts)
+
+data3b[data3b$doby == "1976-1985" & data3b$male == "1.Male", "Ownssb_P50"] / data3b[data3b$doby == "1936-1945" & data3b$male == "1.Male", "Ownssb_P50"]
+
+data3b[data3b$doby == "1976-1985" & data3b$male == "2.Female", "Ownssb_P50"] / data3b[data3b$doby == "1936-1945" & data3b$male == "2.Female", "Ownssb_P50"]
+
+1 - data3b[data3b$doby == "1936-1945" & data3b$male == "2.Female", "Ownssb_P50"] / data3b[data3b$doby == "1936-1945" & data3b$male == "1.Male", "Ownssb_P50"]
+
+1 - data3b[data3b$doby == "1976-1985" & data3b$male == "2.Female", "Ownssb_P50"] / data3b[data3b$doby == "1976-1985" & data3b$male == "1.Male", "Ownssb_P50"]
+```
+
+## Theme 9: Gen Xers and Xennials will pay more taxes in retirement
 
@@
   filter(doby %in% cohorts) %>%
-  mutate(avg_annual_tax_rate = (-1 * ncharts_tax_Mean) / ncharts_income_Mean) %>%
+  mutate(avg_annual_tax_rate = (-1 * pc_tot_tax_Mean) / pc_cen_income_Mean) %>%
   mutate(doby = rename_cohort(doby)) %>%    

### 2019-07-02 draft_charts.Rmd
Close issues #3, #4, #5, and #6. Add .Renviron to the .gitignore.

@@
 
-knitr::opts_chunk$set(fig.width = 6.5, fig.height = 5, fig.path = "figures/", dev = c("png", "cairo_pdf"))
+knitr::opts_chunk$set(fig.width = 6.5, fig.height = 6, fig.path = "figures/", dev = c("png", "cairo_pdf"))
 ```
@@
 # create the output plot
-plotr(titler("A typical Xennial retiree will receive about 30 percent more net income than a typical\nretiree born 40 years earlier"),
+plotr(titler("A typical Xennial retiree will receive about 30 percent more net income than a typical retiree born 40 years earlier", width = 60),
       subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.8, 1.5, 30, 1, 1))
+      noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals, minus payroll taxes, federal and state income taxes, and Medicare surtaxes.
+The analysis computes a per capita measure by dividing total family income by two for married people and assumes that scheduled Social Security benefits are paid."),
+      heights = c(4, 1.5, 30, 1, 1, 4))
 ```
@@
 # create the output plot
-plotr(titler("Over her lifetime, a typical Xennial woman will earn more than twice as much as her\npre-boomer counterpart"),
+plotr(titler("Over her lifetime, a typical Xennial woman will earn more than twice as much as her pre-boomer counterpart"),
       subtitler("Median earnings from ages 25 to 69 (constant 2018 dollars)"),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.6, 2, 30, 1, 1))
+      noter("Estimates are restricted to adults who survive to age 70."), 
+      heights = c(4, 2, 30, 1, 1, 1))
 ```
@@
 # create the output plot
-plotr(titler("Nearly one-third of Xennials will see their living standards fall when they retire"),
+plotr(titler("Nearly one-third of Xennials will see their living standards fall when they retire", width = 61),
       subtitler("Share of adults unable to replace 75 percent of preretirement earnings at age 70"),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2, 2, 30, 1, 1))
+      noter("The analysis compares per capita family income at age 70 to average annual family earnings received from ages 50 to 59, both measured in constant 2018 dollars as adjusted by the change in the consumer price index.
+The income measure at age 70 includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, and the annual income from an annuity valued at 80 percent of financial assets (including retirement accounts).
+The analysis computes a per capita measure by dividing total family income by two for married people and assumes that scheduled Social Security benefits are paid."),
+      heights = c(4, 2, 30, 1, 1, 5))
 ```
@@
 # create the output plot
-plotr(titler("Retirement incomes may fall short for about 4 in 10 Xennials if policymakers\nignore Social Security’s financing gap"),
+plotr(titler("Retirement incomes may fall short for about 4 in 10 Xennials if policymakers ignore Social Security’s financing gap"),
       subtitler("Share of adults unable to replace 75 percent of their preretirement earnings at age 70 "),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(3, 1.5, 30, 1, 1))
+      noter("The analysis compares per capita family income at age 70 to average annual earnings received from ages 50 to 59, both measured in constant 2018 dollars as adjusted by the change in the consumer price index.
+The income measure at age 70 includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, and the annual income from an annuity valued at 80 percent of financial assets (including retirement accounts).
+The analysis computes a per capita measure by dividing total family income by two for married people.", width = 131),
+      heights = c(6.5, 1.5, 30, 1, 1, 5))
 ```
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(1.8, 1, 30, 1, 1))
+      noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals.
+The analysis computes a per capita measure by dividing total family income by two for married people.
+The sample is restricted to Social Security beneficiaries.
+The analysis assumes that scheduled Social Security benefits are paid."),
+      heights = c(4, 1, 30, 1, 1, 4))
 ```
@@
 # create the output plot
-plotr(titler("A 25 percent Social Security benefit cut could leave about 1 in 12 Gen X and Xennial\nretirees in poverty"),
+plotr(titler("A 25 percent Social Security benefit cut could leave about 1 in 12 Gen X and Xennial retirees in poverty"),
       subtitler("Percentage of adults with incomes below the federal poverty level at age 70 "),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.8, 1.5, 30, 1, 1))
+      noter("The analysis compares family income from Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, rent, retirement account withdrawals, and income from coresident family members to the federal poverty level.
+"),
+      heights = c(4, 1.5, 30, 1, 1, 3))
 ```
@@
 # create the output plot
-plotr(titler("Although retirement incomes are projected to rise for people of color, they\nwill continue to lag far behind incomes for non-Hispanic white people"),
+plotr(titler("Although retirement incomes are projected to rise for people of color, they will continue to lag far behind incomes for non-Hispanic white people"),
       subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(3, 1.5, 30, 1, 1))
+      noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals, minus payroll taxes, federal and state income taxes, and Medicare surtaxes.
+The analysis computes a per capita measure by dividing total family income by two for married people.
+The analysis assumes that scheduled Social Security benefits are paid.
+The chart excludes other racial groups because of small sample sizes."),
+      heights = c(6.5, 1.2, 28, 1, 1, 4.5))
 ```
@@
 (data6[data6$doby == "1936-1945" & data6$raceeth == "3.Hispanic" & !is.na(data6$raceeth), "pc_cen_net_income_P50"] - data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]) / data6[data6$doby == "1936-1945" & data6$raceeth == "1.White, Non-Hisp" & !is.na(data6$raceeth), "pc_cen_net_income_P50"]
-
 ```
@@
 # create the output plot
-plotr(titler("For Xennials, women’s median Social Security benefits at age 70 will rise to 85 percent of\nthe median for men"),
+plotr(titler("For Xennials, women’s median Social Security benefits at age 70 will rise to 85 percent of the median for men"),
       subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.6, 2, 30, 1, 1))
+      noter("Analysis assumes that promised Social Security benefit payments will continue."), 
+      heights = c(4, 1, 30, 1, 1, 1))
 ```
@@
       sourcer("DYNASIM4, ID963"),
-      heights = c(2.6, 1, 30, 1, 1))
+      noter("Taxes include payroll taxes, federal and state income taxes, and Medicare surtaxes.
+Property, sales, and excise taxes are excluded.\nThe analysis assumes that current tax rules continue indefinitely and that scheduled Social Security benefits are paid."),
+      heights = c(2, 1, 30, 1, 1, 2))
 ```

### 2019-07-03 draft_charts.Rmd
Add periods to Source lines.

@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals, minus payroll taxes, federal and state income taxes, and Medicare surtaxes.
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("Estimates are restricted to adults who survive to age 70."), 
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("The analysis compares per capita family income at age 70 to average annual family earnings received from ages 50 to 59, both measured in constant 2018 dollars as adjusted by the change in the consumer price index.
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("The analysis compares per capita family income at age 70 to average annual earnings received from ages 50 to 59, both measured in constant 2018 dollars as adjusted by the change in the consumer price index.
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("This income measure includes Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, capital gains, and retirement account withdrawals.
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("The analysis compares family income from Social Security, earnings, defined-benefit pensions, Supplemental Security Income, other government benefits, interest, dividends, rent, retirement account withdrawals, and income from coresident family members to the federal poverty level.
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("Analysis assumes that promised Social Security benefit payments will continue."), 
@@
       caption,
-      sourcer("DYNASIM4, ID963"),
+      sourcer("DYNASIM4, ID963."),
       noter("Taxes include payroll taxes, federal and state income taxes, and Medicare surtaxes.

### 2019-07-12 draft_charts.Rmd

@@
 * Quintiles are based on per-capita, price-adjusted, shared annuitized income with federal taxes, state taxes, payroll taxes and the Medicare surtax at age 70.
-* Values are the mean of the different sources of income in each quintile.
-
-Financial income = per-capita interest + per-capita dividends + per-capita rental income + per-capita IRA withdrawals
-
-Other income = pcNotModel_Mean
 
 ```{r theme5}
-incomes <- c("Other income", "Financial income", "DB pension income", "Earnings", "Social Security")
 quintile <- c("Bottom quintile", "Second quintile", "Third quintile", "Fourth quintile", "Top quintile")

### 2019-07-15 draft_charts.Rmd
Change legend in chart 2 and title in chart 7

@@
     cohort <- recode(x,
-      `1936-1945` = "Pre-boomers\n(1936—1945)", 
-      `1946-1955` = "Early boomers\n(1946—1955)", 
-      `1956-1965` = "Late boomers\n(1956—1965)", 
-      `1966-1975` = "Gen Xers\n(1966—1975)", 
-      `1976-1985` = "Xennials\n(1976—1985)")
+      `1936-1945` = "Pre-boomers\n(1936–1945)", 
+      `1946-1955` = "Early boomers\n(1946–1955)", 
+      `1956-1965` = "Late boomers\n(1956–1965)", 
+      `1966-1975` = "Gen Xers\n(1966–1975)", 
+      `1976-1985` = "Xennials\n(1976–1985)")
 
-    factor(cohort, levels = c("Pre-boomers\n(1936—1945)", 
-                              "Early boomers\n(1946—1955)", 
-                              "Late boomers\n(1956—1965)", 
-                              "Gen Xers\n(1966—1975)", 
-                              "Xennials\n(1976—1985)"))
+    factor(cohort, levels = c("Pre-boomers\n(1936–1945)", 
+                              "Early boomers\n(1946–1955)", 
+                              "Late boomers\n(1956–1965)", 
+                              "Gen Xers\n(1966–1975)", 
+                              "Xennials\n(1976–1985)"))
   } else {
     cohort <- recode(x,
-      `1936-1945` = "Pre-boomers (1936—1945)", 
-      `1946-1955` = "Early boomers (1946—1955)", 
-      `1956-1965` = "Late boomers (1956—1965)", 
-      `1966-1975` = "Gen Xers (1966—1975)", 
-      `1976-1985` = "Xennials (1976—1985)")
+      `1936-1945` = "Pre-boomers (1936–1945)", 
+      `1946-1955` = "Early boomers (1946–1955)", 
+      `1956-1965` = "Late boomers (1956–1965)", 
+      `1966-1975` = "Gen Xers (1966–1975)", 
+      `1976-1985` = "Xennials (1976–1985)")
 
-    factor(cohort, levels = c("Pre-boomers (1936—1945)", 
-                              "Early boomers (1946—1955)", 
-                              "Late boomers (1956—1965)", 
-                              "Gen Xers (1966—1975)", 
-                              "Xennials (1976—1985)"))
+    factor(cohort, levels = c("Pre-boomers (1936–1945)", 
+                              "Early boomers (1946–1955)", 
+                              "Late boomers (1956–1965)", 
+                              "Gen Xers (1966–1975)", 
+                              "Xennials (1976–1985)"))
   }
@@
   mutate(male = recode(male, 
-                       `1.Male` = "Male",
-                       `2.Female` = "Female")) %>%
+                       `1.Male` = "Men",
+                       `2.Female` = "Women")) %>%
   filter(doby %in% cohorts) %>%
@@
 plotr(titler("Retirement incomes may fall short for about 4 in 10 Xennials if policymakers ignore Social Security’s financing gap"),
-      subtitler("Share of adults unable to replace 75 percent of their preretirement earnings at age 70 "),
+      subtitler("Share of adults unable to replace 75 percent of their preretirement earnings at age 70"),
       chart9,
@@
 plotr(titler("A 25 percent Social Security benefit cut could leave about 1 in 12 Gen X and Xennial retirees in poverty"),
-      subtitler("Percentage of adults with incomes below the federal poverty level at age 70 "),
+      subtitler("Share of adults with incomes below the federal poverty level at age 70 "),
       chart12,
@@
 # create the output plot
-plotr(titler("Although retirement incomes are projected to rise for people of color, they will continue to lag far behind incomes for non-Hispanic white people"),
+plotr(titler("Although retirement incomes are projected to rise for people of color, they will continue to lag far behind incomes for white people"),
       subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
@@
   mutate(male = recode(male, 
-                       `1.Male` = "Male",
-                       `2.Female` = "Female")) %>%  
+                       `1.Male` = "Men",
+                       `2.Female` = "Women")) %>%  
   filter(doby %in% cohorts) %>%

### 2019-07-22 draft_charts.Rmd
Update title spacing and add periods to sources.

@@
 # create the output plot
-plotr(titler("A typical Xennial retiree will receive about 30 percent more net income than a typical retiree born 40 years earlier", width = 60),
+plotr(titler("A typical Xennial retiree will receive about 24 percent more net income than a typical retiree born 40 years earlier", width = 60),
       subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
@@
 The analysis computes a per capita measure by dividing total family income by two for married people and assumes that scheduled Social Security benefits are paid."),
-      heights = c(4, 1.5, 30, 1, 1, 4))
+      heights = c(3.2, 1.5, 30, 1, 1, 4))
 ```
@@
 # create the output plot
-plotr(titler("Over her lifetime, a typical Xennial woman will earn more than twice as much as her pre-boomer counterpart"),
+plotr(titler("Over her lifetime, a typical Xennial woman will earn more\nthan twice as much as her pre-boomer counterpart", width = FALSE),
       subtitler("Median earnings from ages 25 to 69 (constant 2018 dollars)"),
@@
       noter("Estimates are restricted to adults who survive to age 70."), 
-      heights = c(4, 2, 30, 1, 1, 1))
+      heights = c(3.2, 1.5, 30, 1, 1, 1))
 ```
@@
 # create the output plot
-plotr(titler("Nearly one-third of Xennials will see their living standards fall when they retire", width = 61),
+plotr(titler("Nearly one-third of Xennials will see their\nliving standards fall when they retire", width = FALSE),
       subtitler("Share of adults unable to replace 75 percent of preretirement earnings at age 70"),
@@
 The analysis computes a per capita measure by dividing total family income by two for married people.", width = 131),
-      heights = c(6.5, 1.5, 30, 1, 1, 5))
+      heights = c(4, 1.8, 30, 1, 1, 5))
 ```
@@
 # create the output plot
-plotr(titler("Social Security accounts for about 85 percent of income for low-income beneficiaries"),
+plotr(titler("Social Security accounts for about 85 percent of\nincome for low-income beneficiaries", width = FALSE),
       subtitler("Share of per capita family income at age 70 from Social Security by income quintile"),
@@
 The analysis assumes that scheduled Social Security benefits are paid."),
-      heights = c(4, 1, 30, 1, 1, 4))
+      heights = c(3.5, 1.3, 30, 1, 1, 4))
 ```
@@
 # create the output plot
-plotr(titler("A 25 percent Social Security benefit cut could leave about 1 in 12 Gen X and Xennial retirees in poverty"),
+plotr(titler("A 25 percent Social Security benefit cut could leave about\n1 in 12 Gen X and Xennial retirees in poverty",width = FALSE),
       subtitler("Share of adults with incomes below the federal poverty level at age 70 "),
@@
 "),
-      heights = c(4, 1.5, 30, 1, 1, 3))
+      heights = c(3.5, 1.3, 30, 1, 1, 3))
 ```
@@
 The chart excludes other racial groups because of small sample sizes."),
-      heights = c(6.5, 1.2, 28, 1, 1, 4.5))
+      heights = c(3.5, 1.2, 28, 1, 1, 4.5))
 ```
@@
 # create the output plot
-plotr(titler("For Xennials, women’s median Social Security benefits at age 70 will rise to 85 percent of the median for men"),
+plotr(titler("For Xennials, women’s median Social Security benefits at\nage 70 will rise to 85 percent of the median for men", width = FALSE),
       subtitler("Median annual Social Security income at age 70 (constant 2018 dollars)"),
@@
       noter("Analysis assumes that promised Social Security benefit payments will continue."), 
-      heights = c(4, 1, 30, 1, 1, 1))
+      heights = c(3.5, 1.5, 30, 1, 1, 1))
 ```
@@
 Property, sales, and excise taxes are excluded.\nThe analysis assumes that current tax rules continue indefinitely and that scheduled Social Security benefits are paid."),
-      heights = c(2, 1, 30, 1, 1, 2))
+      heights = c(1.5, 1, 30, 1, 1, 2))
 ```

### 2019-07-22 urban_style.R

@@ -1,4 +1,4 @@
-titler <- function(title, width = 58) {
+titler <- function(title, width = 68) {
   
   if (width == FALSE) {
     textGrob(title, 
@@ -6,18 +6,20 @@ titler <- function(title, width = 58) {
              y = unit(1, "npc"), 
              hjust = 0, 
              vjust = 1,
-             gp = gpar(fontsize = 18, 
+             gp = gpar(fontsize = 14, 
                        fontfamily = "Lato", 
-                       lineheight = 1))    
+                       lineheight = 1,
+                       fontface = "bold"))    
   } else {
     textGrob(stringr::str_wrap(title, width = width), 
              x = unit(0, "npc"), 
              y = unit(1, "npc"), 
              hjust = 0, 
              vjust = 1,
-             gp = gpar(fontsize = 18, 
+             gp = gpar(fontsize = 14, 
                        fontfamily = "Lato", 
-                       lineheight = 1))
+                       lineheight = 1,
+                       fontface = "bold"))
   }
 }
 
@@ -27,7 +29,7 @@ subtitler <- function(subtitle) {
            y = unit(1, "npc"),
            hjust = 0, 
            vjust = 1,
-           gp = gpar(fontsize = 12, 
+           gp = gpar(fontsize = 10, 
                      fontfamily = "Lato"))
 }
 

### 2019-07-22 draft_charts.Rmd
Drop about from title 1 and the note in chart 8.

@@
 # create the output plot
-plotr(titler("A typical Xennial retiree will receive about 24 percent more net income than a typical retiree born 40 years earlier", width = 60),
+plotr(titler("A typical Xennial retiree will receiv 24 percent more net income\nthan a typical retiree born 40 years earlier", width = FALSE),
       subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
@@
       sourcer("DYNASIM4, ID963."),
-      noter("Analysis assumes that promised Social Security benefit payments will continue."), 
+      noter("The analysis assumes that promised Social Security benefit payments will continue."), 
       heights = c(3.5, 1.5, 30, 1, 1, 1))

### 2019-07-22 draft_charts.Rmd
Fix typo in chart 1 title and adjust line breaks in titles for charts 3 and 5

@@
 # create the output plot
-plotr(titler("A typical Xennial retiree will receiv 24 percent more net income\nthan a typical retiree born 40 years earlier", width = FALSE),
+plotr(titler("A typical Xennial retiree will receive 24 percent more net income\nthan a typical retiree born 40 years earlier", width = FALSE),
       subtitler("Median annual per capita after-tax family income at age 70 (constant 2018 dollars)"),
@@
 # create the output plot
-plotr(titler("Nearly one-third of Xennials will see their\nliving standards fall when they retire", width = FALSE),
+plotr(titler("Nearly one-third of Xennials will see their living standards\nfall when they retire", width = FALSE),
       subtitler("Share of adults unable to replace 75 percent of preretirement earnings at age 70"),
@@
 # create the output plot
-plotr(titler("Social Security accounts for about 85 percent of\nincome for low-income beneficiaries", width = FALSE),
+plotr(titler("Social Security accounts for about 85 percent of income\nfor low-income beneficiaries", width = FALSE),
       subtitler("Share of per capita family income at age 70 from Social Security by income quintile"),
